---
title: "ArchiGWAS"
author: "Opamelas83"
date: "2023-05-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
```{r}
install.packages("devtools"); library(devtools)
install.packages("sommer") #new version is important
library(tidyverse); library(data.table); library(genomicMateSelectR); library(magrittr); library(sommer); library(corrplot); library(qqman)
install.packages("devtools")
#devtools::install_github("jiabowang/GAPIT3",force=TRUE) not working for me
#library(GAPIT)
source("http://zzlab.net/GAPIT/gapit_functions.txt")
#library(snpsOfInterest)
#str(snpsOfInterest) It is possible to color the important SNP= SNP of unterest.
```
#PART 1 — Prepare the 3,113 accession list for PLINK
#1. PREPARE ACCESSION LIST FOR LD ANALYSIS
```{r}
GWAS_ID <- readRDS(
  here::here("output", "GWAS_accession_IDs.rds"))
GWAS_ID_PLINK <- data.frame(FID = GWAS_ID, IID = GWAS_ID)
write.table(GWAS_ID_PLINK,
  here::here("output", "GWAS_accession_IDs_PLINK.txt"),
  row.names = FALSE, col.names = FALSE, quote = FALSE, sep = "\t")
nrow(GWAS_ID_PLINK)# 3113

```
#PART 2 — My PLINK LD analysis
#In Terminal
```{r}
#mo367@Michelle-okoma ~ % cd /Users/mo367/Desktop/Cornell/Analy_project/Plantarchitect/data/VCF
#mo367@Michelle-okoma VCF % plink --vcf BreedBaseGenotypes.vcf.gz \
#      --double-id \
#      --allow-extra-chr \
#      --make-bed \
#      --out cassava
##to restrict the LD analysis to your 3,113 accessions
#plink --bfile cassava \
#      --keep ../../output/GWAS_accession_IDs_PLINK.txt \
#      --r2 gz \
#      --ld-window 999999 \
#      --ld-window-kb 1000 \
#      --ld-window-r2 0 \
#      --out cassava_LD_GWAS3113
```
#In terminal how does the whole file look like 
```{r}
#PLINK v1.90b7 64-bit (16 Jan 2023)             www.cog-genomics.org/plink/1.9/
#(C) 2005-2023 Shaun Purcell, Christopher Chang   GNU General Public License v3
#Logging to cassava.log.
#Options in effect:
#  --allow-extra-chr
#  --double-id
#  --make-bed
#  --out cassava
#  --vcf BreedBaseGenotypes.vcf.gz

#16384 MB RAM detected; reserving 8192 MB for main workspace.
#--vcf: cassava-temporary.bed + cassava-temporary.bim + cassava-temporary.fam
#written.
#61239 variants loaded from .bim file.
#7417 people (0 males, 0 females, 7417 ambiguous) loaded from .fam.
#Ambiguous sex IDs written to cassava.nosex .
#Using 1 thread (no multithreaded calculations invoked).
#Before main variant filters, 7417 founders and 0 nonfounders present.
#Calculating allele frequencies... done.
#Total genotyping rate is exactly 1.
#61239 variants and 7417 people pass filters and QC.
#Note: No phenotypes present.
#--make-bed to cassava.bed + cassava.bim + cassava.fam ... done.
#mo367@Michelle-okoma VCF % plink --bfile cassava \
#      --r2 gz \
#      --ld-window 999999 \
#      --ld-window-kb 1000 \
#      --ld-window-r2 0 \
#      --out cassava_LD
#PLINK v1.90b7 64-bit (16 Jan 2023)             #www.cog-genomics.org/plink/1.9/
#(C) 2005-2023 Shaun Purcell, Christopher Chang   GNU General Public #License v3
#Logging to cassava_LD.log.
#Options in effect:
# --bfile cassava
# --ld-window 999999
# --ld-window-kb 1000
# --ld-window-r2 0
# --out cassava_LD
#  --r2 gz

#16384 MB RAM detected; reserving 8192 MB for main workspace.
#61239 variants loaded from .bim file.
#7417 people (0 males, 0 females, 7417 ambiguous) loaded from .fam.
#Ambiguous sex IDs written to cassava_LD.nosex .
#Using up to 9 threads (change this with --threads).
#Before main variant filters, 7417 founders and 0 nonfounders present.
#Calculating allele frequencies... done.
#Total genotyping rate is exactly 1.
#61239 variants and 7417 people pass filters and QC.
#Note: No phenotypes present.
#--r2 to cassava_LD.ld.gz ... done.          
#mo367@Michelle-okoma VCF % ls cassava_LD*
#cassava_LD.ld.gz	cassava_LD.log		cassava_LD.nosex
```
#PART 3 — Read the new LD results in R
```{r}
## 2. LINKAGE DISEQUILIBRIUM DECAY
#install.packages('R.utils')
library(R.utils)
library(data.table)
library(data.table)
library(dplyr)
library(ggplot2)
library(scales)
# Read PLINK LD output (gz OK)
LDfile <- fread("data/VCF/cassava_LD_GWAS3113.ld.gz")#expects columns incl. BP_A, BP_B, R2)
# Calculate physical distance between SNP pairs
LDfile <- LDfile %>% mutate(DIST_BP = abs(BP_B - BP_A)) %>% 
  filter(DIST_BP > 0, R2 < 1)
# Bin pairwise physical distances
breaks <- unique(floor(10^seq(1, 6, by = 0.1)))   # ~10 bp to 1,000,000 bp
LDfile$BIN <- cut(LDfile$DIST_BP, breaks = c(0, breaks), 
                  include.lowest = TRUE)
# Mean LD within each distance bin
decay <- LDfile %>%
  group_by(BIN) %>%
  summarise(DIST_MID = median(DIST_BP), R2_MEAN = mean(R2), N = n(), .groups="drop") %>%
  arrange(DIST_MID)
# First distance bin at which mean r² reaches ≤ 0.20
d_at_02 <- decay %>%
  filter(!is.na(R2_MEAN), R2_MEAN <= 0.2) %>%
  arrange(DIST_MID) %>%
  slice(1) %>%
  pull(DIST_MID)
cat(sprintf("Mean r^2 ~ 0.2 at ~ %.0f kb\n", d_at_02/1000))
# Add the mean LD value at 1,000 kb for visualization
last_distance <- max(decay$DIST_MID, na.rm = TRUE)

R2_1000 <- LDfile %>%
  filter(
    DIST_BP > last_distance,
    DIST_BP <= 1000000
  ) %>%
  summarise(
    R2_MEAN = mean(R2, na.rm = TRUE),
    N = n()
  )

decay_plot <- decay %>%
  dplyr::select(DIST_MID, R2_MEAN, N) %>%
  bind_rows(
    data.frame(
      DIST_MID = 1000000,
      R2_MEAN = R2_1000$R2_MEAN,
      N = R2_1000$N
    )
  ) %>%
  arrange(DIST_MID)
# Sample 200,000 SNP pairs for visualization only
LDplot <- LDfile %>%
  filter(
    DIST_BP > 0,
    DIST_BP <= 1000000,
    R2 < 1
  )

set.seed(123)

if (nrow(LDplot) > 200000) {
  LDplot <- slice_sample(LDplot, n = 200000)
}
```
#FINAL GENOME-WIDE LD-decay figure PLOT FIGURE 8
```{r}
# 5) Plot LD decay (publication-ready)
p <- ggplot() +
# Individual pairwise LD estimates; 200,000 pairs sampled for visualization only
geom_point(data = LDplot, aes(x = DIST_BP / 1000, y = R2),
color = "grey40", size = 0.30, alpha = 0.08) +
# Mean r² for each physical-distance bin
geom_point(data = decay_plot,
aes(x = DIST_MID / 1000, y = R2_MEAN),
color = "red", size = 2.4) +
# Connect the mean r² values
geom_line(data = decay_plot, aes(x = DIST_MID / 1000, y = R2_MEAN),
color = "red", linewidth = 1.2) +
scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 200)) +
scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
labs(x = "Distance between SNPs (kb)", y = expression(r^2)) +
theme_classic(base_size = 14)
p
ggsave(filename = "Result/Figure8_LD_decay.pdf",
  plot = p, width = 8, height = 7, units = "in")
```
## PREPARE PHENOTYPE DATA FOR GAPIT
```{r}
blups <- readRDS(here::here("output", "blups.rds"))

GWAS_ID <- readRDS(
  here::here("output", "GWAS_accession_IDs.rds")
)

blup1 <- blups[[2]][[1]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(BranchlevelNum = BLUP) %>%
  dplyr::select(Taxa, BranchlevelNum)

blup2 <- blups[[2]][[2]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(FirstBranchHeight = BLUP) %>%
  dplyr::select(Taxa, FirstBranchHeight) %>%
  right_join(blup1, by = "Taxa")

blup3 <- blups[[2]][[3]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(PlantHeight = BLUP) %>%
  dplyr::select(Taxa, PlantHeight) %>%
  right_join(blup2, by = "Taxa")

blup4 <- blups[[2]][[4]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(Cylindrical = BLUP) %>%
  dplyr::select(Taxa, Cylindrical) %>%
  right_join(blup3, by = "Taxa")

blup5 <- blups[[2]][[5]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(Umbrella = BLUP) %>%
  dplyr::select(Taxa, Umbrella) %>%
  right_join(blup4, by = "Taxa")

blup6 <- blups[[2]][[6]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(Open = BLUP) %>%
  dplyr::select(Taxa, Open) %>%
  right_join(blup5, by = "Taxa")

blupT <- blups[[2]][[7]] %>%
  rename(Taxa = GermplasmName) %>%
  mutate(Compact = BLUP) %>%
  dplyr::select(Taxa, Compact) %>%
  right_join(blup6, by = "Taxa")

# Restrict phenotype table explicitly to final GWAS accessions
blupT <- blupT %>%
  filter(Taxa %in% GWAS_ID)

write.table(
  blupT,
  here::here("data", "MyArchiphenos_blupPP"),
  row.names = FALSE,
  quote = FALSE
)
```
#GAPIT included in manuscript NEW VERSION
```{r}
#Step 1: Set data directory and import files
myY <- read.table("data/MyArchiphenos_blupPP", head = TRUE)
myG <- read.table("data/VCF/BreedBaseGenotypes.hmp.txt", head = FALSE)#
 dim(myY)#: 3113    8
 length(unique(myY$Taxa))#: 3113
 dim(myG)#: 61239  7428

##For running GAPIT in R new version
source("https://raw.githubusercontent.com/jiabowang/GAPIT/refs/heads/master/gapit_functions.txt", encoding = "UTF-8")
 
#Step 2: Run GAPIT
myGAPIT <- GAPIT(
Y=myY,
G=myG,
PCA.total=3,
model=c("BLINK","MLM"),
SNP.MAF = 0.01)#,
# Run single GAPIT multi-method figure for BranchlevelNum to show the shared chromosome 2 signal #between MLM and BLINK rather than presenting the two Manhattan plots separately.and FirstBranchHeight phenotype only 

myY_Branch <- myY %>%
  dplyr::select(Taxa, BranchlevelNum)

myGAPIT <- GAPIT(
Y=myY_Branch,
G=myG,
PCA.total=3,
model=c("BLINK","MLM"),
SNP.MAF = 0.01)
#Do the same for each trait for the remain traits in the suplementary file.
myY_Remain <- myY %>%
  dplyr::select(Taxa, PlantHeight, Compact, Open, Umbrella, Cylindrical)

myGAPIT <- GAPIT(
Y=myY_Remain,
G=myG,
PCA.total=3,
model=c("BLINK","MLM"),
SNP.MAF = 0.01)

## For genomic selection
#myGAPIT <- GAPIT(
#Y=myY,
#G=myG,
#PCA.total=3,
#model= c("gBLUP"),
#SNP.MAF = 0.01)
#inf.pred=merge(myGAPIT$Pred[myGAPIT$Pred[,3]==2,c(1,3,5,8)],myY,by.x="Taxa",by.y="Taxa")
#cor(inf.pred[,4],inf.pred[,5])
```
### SUPPLEMENTARY TABLE S2: Pairwise LD among significant chromosome 2 SNPs associated with BranchlevelNum
```{r}
#Read PLINK LD output (gz OK)
LDfile <- fread("data/VCF/cassava_LD_GWAS3113.ld.gz")#expects columns incl. BP_A, BP_B, R2)
# Significant chromosome 2 SNPs from the revised MLM GWAS
BranchlevelNum_chr2_SNPs <- c(
  "S2_2809137",
  "S2_2818574",
  "S2_2828773",
  "S2_2846222",
  "S2_2854244",
  "S2_3000720")
# Extract pairwise LD comparisons among these SNPs
Table_S2 <- LDfile %>%
  filter(
    CHR_A == 2,
    CHR_B == 2,
    SNP_A %in% BranchlevelNum_chr2_SNPs,
    SNP_B %in% BranchlevelNum_chr2_SNPs
  ) %>%
  dplyr::select(
    CHR_A, BP_A, SNP_A,
    CHR_B, BP_B, SNP_B,
    R2
  ) %>%
  arrange(desc(R2))#15
write.csv(
  Table_S2,
  "output/Table_S2_Chr2_BranchlevelNum_LD.csv",
  row.names = FALSE
)
```

