---
title: "Archi_scale"
author: "Opamelas83"
date: "2023-05-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Distribution of plant shape
```{r}
library(tidyverse); library(data.table); library(here); library(magrittr); library(lubridate); library(MuMIn); library(arm); library(lemon); library(reactablefmtr); library(corrplot);library(gridExtra); library(grid); library(gridtext) 
MyArchiphenos <- read_csv("data/MyArchiphenotypes.csv", col_names = TRUE)
ShapeTRaits <- c("Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type")
Shapephenos <- MyArchiphenos %>% 
     # Convert the data to "long format" 
     pivot_longer(cols = all_of(ShapeTRaits), 
                  names_to = "ShapeTRait", values_to = "ShapeValue") %>%
     # Remove missing values 
     filter(!is.na(ShapeValue)) %>% 
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName))

Shapephenos_plant <- Shapephenos %>% dplyr::select(Year, trialType, ShapeTRait, ShapeValue) %>% filter (!is.na(ShapeValue)) %>% mutate(`Plant type` = ShapeValue)
#ShapeTRait", values_to = "ShapeValue"
Shapephenos_plantsum <- Shapephenos_plant %>% filter(!(`Plant type`) %in% c("Non_Cylindrical", "Non_Umbrella", "Non_Open","Non_Compact")) %>% group_by(trialType, `Plant type`) %>% summarise(numb = length(`Plant type`)) %>% mutate(`Value (%)` = round((numb/rowsum(numb, trialType)) * 100)) #"Non_Erect"

Shapephenos_plantsum$trialType <- factor(Shapephenos_plantsum$trialType, 
                                      levels = c("CET", "PYT", "AYT", "UYT"))
Shapephenos_plantsum$`plant shape` <- factor(Shapephenos_plantsum$`Plant type`,      
                        levels = c("Compact", "Open", "Umbrella", "Cylindrical"))
#jpeg(filename = "Result/PlantShapedistri.jpeg")
ggplot(data = Shapephenos_plantsum, aes(x = trialType, y = `Value (%)`, 
                                group = `plant shape`)) +
  geom_point(aes(color = `plant shape`), size = 3) +
  geom_line(aes(color=`plant shape`), linetype = 81) +
  scale_color_brewer(palette = "Dark2") +
 ggtitle("Plant shape") +
  xlab("breeding stage") #+
ggsave("Result/PlantShapedistri.png",  scale = 1,)
#theme(axis.title.x=element_blank())
#dev.off()
```
### Field trials analysis: Run the model for Plantshape replicated trials
```{r}
MyArchiphenos <- read_csv("data/MyArchiphenotypes.csv", col_names = TRUE)
ShapeTRaits <- c("Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type")
Shapephenos <- MyArchiphenos %>% 
     # Convert the data to "long format" 
     pivot_longer(cols = all_of(ShapeTRaits), 
                  names_to = "ShapeTRait", values_to = "ShapeValue") %>%
     # Remove missing values 
     filter(!is.na(ShapeValue)) %>% 
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName))

ShapeTRials <- Shapephenos %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names
i <- ShapeTRaits [1]
j <- ShapeTRials[1]
results_Shape<- tibble()
for(i in ShapeTRaits){
 for(j in ShapeTRials){
     Data <- Shapephenos %>% 
       filter(studyName == j, ShapeTRait == i) %>% 
     filter(!is.na(germplasmName))
     if(nrow(Data) > 1){
      if(length(unique(Data$ShapeValue)) > 1){ 
     try(MixedModels_Shape <- glmer(as.factor(ShapeValue) ~ 1 + (1|germplasmName), data = Data,  family = binomial, nAGQ = 2))#repInTrial +
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trait = i,
                      Trial = j,
                      VarG = as.data.frame(VarCorr(MixedModels_Shape))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(var(residuals(MixedModels_Shape))) %>% .[1,1],
                      Mean = as.data.frame(fixef(MixedModels_Shape))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),2),
                      CV = sigma(MixedModels_Shape)/Mean,
                      PEV = mean((se.ranef(MixedModels_Shape)$germplasmName[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),2)))
     try(results_Shape <- bind_rows(results_Shape, result))
     #rm(MixedModels_Shape); rm(result)
      }
  }   
}
}
#results_Shape
#var(residuals(MixedModels_Shape))
#[1] 0.006056236
#var(summary()$residuals)
#[1] 0.003159229
write.table(results_Shape, file = "Result/results_Shape.csv", quote = T, sep =",", append = FALSE, row.names = F)

write.table(Shapephenos, file = "data/Shapephenos.csv", quote = T, sep =",", append = FALSE, row.names = F)
```
# I would not filter for shape traits because the heritability was mostly high. but it still improve the general heritability. "(trials Filter trials for identify trials with low heritability (H2 < 0.1) and/or low accuracy (Ac < 0.4) for all the traits)"
```{r}
#Filter with H2 and Ac
results_Shape <- read_csv("Result/results_Shape.csv", col_names = TRUE)
# Determine the limit values of H2 and Ac by trait 
#results_Shape %>% summary()
# Visualize the distribution 
#plot(results_Shape$H2, results_Shape$Ac) # It is a challenge to chose the value of H2 and AC. Here the trials have high values of both.
Trials_to_keep <- results_Shape %>% filter((!is.na(H2) & H2 > 0.1) | (!is.na(Ac) & Ac >= 0.4)) # I should remove the trials by trait since a given trial can be in the "remove position" for 1 trait but be in the "stay position" for another trait. 
ShapeTRaits <- c("Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type", "Erectibility")#

Shapephenos_filtered <- tibble()
for(i in ShapeTRaits){
  subdata <- Trials_to_keep %>% filter(Trait == i) 
  subdata <- subdata$Trial
  Shapephenos_filt <- Shapephenos %>% filter(ShapeTRait == i) 
  Shapephenos_filtered1 <- Shapephenos_filt %>% dplyr::filter(studyName %in% subdata)
 Shapephenos_filtered <- bind_rows(Shapephenos_filtered, Shapephenos_filtered1)
}
write.table(Shapephenos_filtered, file = "data/Shapephenos_filtered.csv", quote = TRUE, sep =",", append = FALSE, row.names = F)
```



