---
title: "fieldtrialanalysis"
author: "Opamelas83"
date: "2023-02-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## check the number of each trait observations by trial. because most of the trials have less than the 3 traits observed. 
```{r}
library(tidyverse); library(data.table); library(lme4); library(here); library(magrittr); library(lubridate); library(MuMIn); library(arm)
MyArchiphenos <- read.csv("data/MyArchiphenotypes.csv")
#The pipeline version of this analysis will use the TRUE/FALSE values of CompleteBlocks and IncompleteBlocks (Preliminary analysis of trial data ).
MyArchiphenos  %>% 
     count(Replicated,non_Replicated,Location) %>% 
     spread(Location,n)
```

#Convert the data to “long format” . Remove missing values. “Nest” the data by Trait.
```{r}
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "PlantArchitecture")
MyArchiphenos <- MyArchiphenos %>% 
     # Convert the data to "long format" 
     pivot_longer(cols = all_of(ArchiTRaits), 
                  names_to = "ArchiTRait", values_to = "ArchiValue") %>%
     # Remove missing values 
     filter(!is.na(ArchiValue)) %>% 
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName))
```

##Fit models for replicated trials 
```{r}
MyArchiphenos_replicated <- MyArchiphenos %>% filter(Replicated == TRUE)
To_remove <- MyArchiphenos_replicated  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_replicate=length(unique(replicate))) %>%
  filter(N_unique_accession == 1 | N_unique_replicate == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_replicated <- MyArchiphenos_replicated %>% dplyr::filter(!studyName %in% Toremove)#this MyArchi.. took to remove outliers
ArchiTRials <- MyArchiphenos_replicated$studyName %>% table %>% .[order(.)] %>% names
 
i <- ArchiTRaits[1]
j <- ArchiTRials[1]
results <- tibble()
results_stu <- tibble()
for(i in ArchiTRaits){
 for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated %>% 
       filter(studyName == j, ArchiTRait == i)
     if(nrow(Data) > 1){
     try(MixedModels <- lmer(ArchiValue ~ repInTrial+ (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trait = i,
                      Trial = j,
                          VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2],
                      
                      Mean = as.data.frame(fixef(MixedModels))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),1),
                      CV = sigma(MixedModels)/Mean,
                      PEV = mean((se.ranef(MixedModels)$germplasmName[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),1)))
 ##Calculate Studentized Residuals
 try(stud_resids <- tibble(StResiduals = stats::rstudent(MixedModels),
     germplasmName = Data$germplasmName))
 ##Bind all results
     try(results <- bind_rows(results, result))
 try(results_stu <- bind_rows(results_stu, stud_resids))
     rm(MixedModels); rm(result); rm(stud_resids)
     }
 }
 }
results
results_stu
write.table(results, file = "Result/MixedModel_replicated.csv", quote = F, sep =",",
            row.names = F)
write.table(results_stu, file = "data/St_residuall_replicated.csv", quote = F, sep =",",
            row.names = F)
```

# Filter trials for identify trials with low heritability (H2 < 0.1) and/or low accuracy (Ac < 0.4) for all the traits, then, removes the outliers with residual value > 3.
```{r}
#add studentized residuals to orignal dataset
results_stu <- read.csv("data/St_residuall_replicated.csv")
MyArchiphenos_replicated_St <- cbind(MyArchiphenos_replicated, results_stu['StResiduals'])
write.table(MyArchiphenos_replicated_St, file = "data/MyArchiphenos_replicated_St.csv", quote = F, sep =",", row.names = F)
#Filter with H2 and Ac
results <- read.csv("Result/MixedModel_replicated.csv")
# Visualize the distribution 
results <- results %>% filter(!is.na(H2) & H2 < 1 & H2 != 0) %>% filter(!is.na(Ac) & Ac < 1 & H2 != 0)#VarG/H2 > 1 for high replicate level (> accession number? but = 6,10,9 ...)
plot(results$H2, results$Ac) # It is a challenge to chose the value of H2 and AC
Trials_replicated_to_remove <- results %>% filter(H2 <= 0.1 | H2 >= 1 | Ac <= 0.4) %>% dplyr::select(Trial)#Should I remove the trials by trait? since a given trial can be in the "remove position" for 1 trait but be in the "stay position" for another trait. 
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "PlantArchitecture")
# Make a list of trial to remove per trait
for(i in ArchiTRaits){
  subdata <- results %>% filter(Trait == i) %>% 
  filter(H2 <= 0.1 | H2 >= 1 | Ac <= 0.4) %>% dplyr::select(Trial)
   paste(ArchiTRaits[i] <- subdata)}

Listfor_BranchlevelNum <- ArchiTRaits$BranchlevelNum
remove_for_BranchlevelNum <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "BranchlevelNum") %>% dplyr::filter(!studyName %in% Listfor_BranchlevelNum)

Listfor_FirstBranchHeight <- ArchiTRaits$FirstBranchHeight
remove_for_FirstBranchHeight <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "FirstBranchHeight") %>% dplyr::filter(!studyName %in% Listfor_FirstBranchHeight)

Listfor_PlantHeight <- ArchiTRaits$PlantHeight
remove_for_PlantHeight <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "PlantHeight") %>% dplyr::filter(!studyName %in% Listfor_PlantHeight)

Listfor_PlantArchitecture <- ArchiTRaits$PlantArchitecture
remove_for_PlantArchitecture <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "PlantArchitecture") %>% dplyr::filter(!studyName %in% Listfor_PlantArchitecture)

MyArchiphenos_replicated_filt <- bind_rows(remove_for_BranchlevelNum, remove_for_FirstBranchHeight, remove_for_PlantHeight, remove_for_PlantArchitecture)

# remove the outliers 
plot(MyArchiphenos_replicated_filt$StResiduals)
outliers <- which(abs(scale(MyArchiphenos_replicated_filt$StResiduals)) > 4)
MyArchiphenos_replicated_filtered <- MyArchiphenos_replicated_filt[-outliers,]
write.table(MyArchiphenos_replicated_filtered, file = "data/MyArchiphenos_replicated_filtered.csv", quote = F, sep =",",row.names = F)
#plot(MyArchiphenos_replicated_filtered$StResiduals)
```

##Fit models for the non_replicated trials: Fit a model only with the check, thus the replication will be the block
```{r}
MyArchiphenos_non_rep <- MyArchiphenos %>% filter(Replicated == FALSE) %>% filter(entryType == "check")
To_remove <- MyArchiphenos_non_rep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% Toremove)
#germplasmtoremove <- MyArchiphenos_non_rep  %>% group_by(studyName, germplasmName) %>% summarize(N_observation = length(germplasmName)) %>% filter(N_observation == 1) #to check the trials including germplasm with 1 value
#germplasmtoremove_trials <- germplasmtoremove$studyName
#germplasmtoremove_germ <- germplasmtoremove$germplasmName
#MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% filter((!studyName %in% germplasmtoremove_trials) & (!germplasmName %in% germplasmtoremove_germ))
 
ArchiTRials <- MyArchiphenos_non_rep %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

i <- ArchiTRaits[1]
j <- ArchiTRials[1]
results <- tibble()
results_stu <- tibble()
for(i in ArchiTRaits){
 for(j in ArchiTRials){
     Data <- MyArchiphenos_non_rep %>% 
       filter(studyName == j, ArchiTRait == i) %>%
       filter(!is.na(ArchiValue))
     if(nrow(Data) > 1){
     MixedModels <- lmer(ArchiValue ~ blockInrep+ (1|germplasmName), data = Data)
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 result <- tibble(Trait = i,
                      Trial = j,
                          VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2],
                      Mean = as.data.frame(fixef(MixedModels))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),1),
                      CV = sigma(MixedModels)/Mean,
                      PEV = mean((se.ranef(MixedModels)$germplasmName[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),1)
 )
##Calculate Studentized Residuals
 stud_resids <- tibble(StResiduals = stats::rstudent(MixedModels),
     germplasmName = Data$germplasmName)
 ##Bind all results
 results <- bind_rows(results, result)
 results_stu <- bind_rows(results_stu, stud_resids)
     #rm(MixedModels); rm(result); rm(stud_resids)
     }
 }
 }
results
results_stu
write.table(results, file = "Result/MixedModel_non_replicated.csv", quote = F, sep =",", row.names = F)
write.table(results_stu, file = "data/St_residual_non_replicated.csv", quote = F, sep =",", row.names = F)
#Error : number of levels of each grouping factor must be < number of observations (problems: germplasmName)
#Error in VarCorr(MixedModels) : object 'MixedModels' not found and Error in list2(...) : object 'result' not found are for are for the where the traits are not measured
#VarG > 1 for high blocs level (> accession number? but = 6,10,9 ...)
```

# Filter trials for identify trials with low heritability (H2 < 0.1) and/or low accuracy (Ac < 0.4) for all the traits
```{r}
#add studentized residuals to orignal dataset: I wwill not remove outliers since we worked with only the checks
#results_stu <- read.csv("data/St_residuall_replicated.csv")
#MyArchiphenos_replicated_St <- cbind(MyArchiphenos_replicated, results_stu['StResiduals'])
#write.table(MyArchiphenos_replicated_St, file = "data/MyArchiphenos_replicated_St.csv", quote = F, sep =",", row.names = F)
#Filter with H2 and Ac
results <- read.csv("Result/MixedModel_replicated.csv")
# Visualize the distribution 
results <- results %>% filter(!is.na(H2) & H2 < 1 & H2 != 0) %>% filter(!is.na(Ac) & Ac < 1 & H2 != 0)#VarG/H2 > 1 for high replicate level (> accession number? but = 6,10,9 ...)
plot(results$H2, results$Ac) # It is a challenge to chose the value of H2 and AC
Trials_replicated_to_remove <- results %>% filter(H2 <= 0.1 | H2 >= 1 | Ac <= 0.4) %>% dplyr::select(Trial)#Should I remove the trials by trait? since a given trial can be in the "remove position" for 1 trait but be in the "stay position" for another trait. 
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "PlantArchitecture")
# Make a list of trial to remove per trait
for(i in ArchiTRaits){
  subdata <- results %>% filter(Trait == i) %>% 
  filter(H2 <= 0.1 | H2 >= 1 | Ac <= 0.4) %>% dplyr::select(Trial)
   paste(ArchiTRaits[i] <- subdata)}

Listfor_BranchlevelNum <- ArchiTRaits$BranchlevelNum
remove_for_BranchlevelNum <- MyArchiphenos_non_rep %>% dplyr::filter(ArchiTRait == "BranchlevelNum") %>% dplyr::filter(!studyName %in% Listfor_BranchlevelNum)

Listfor_FirstBranchHeight <- ArchiTRaits$FirstBranchHeight
remove_for_FirstBranchHeight <-MyArchiphenos_non_rep %>% dplyr::filter(ArchiTRait == "FirstBranchHeight") %>% dplyr::filter(!studyName %in% Listfor_FirstBranchHeight)

Listfor_PlantHeight <- ArchiTRaits$PlantHeight
remove_for_PlantHeight <- MyArchiphenos_non_rep %>% dplyr::filter(ArchiTRait == "PlantHeight") %>% dplyr::filter(!studyName %in% Listfor_PlantHeight)

Listfor_PlantArchitecture <- ArchiTRaits$PlantArchitecture
remove_for_PlantArchitecture <- MyArchiphenos_non_rep %>% dplyr::filter(ArchiTRait == "PlantArchitecture") %>% dplyr::filter(!studyName %in% Listfor_PlantArchitecture)
#I will not forget the trials(remove) for those they were not fited.I should do that by trait. I will skip it because it is only 31 rows(trials?15,...)
#List <- unique(results$Trial)
#remove <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% List)

MyArchiphenos_replicated_filt <- bind_rows(remove_for_BranchlevelNum, remove_for_FirstBranchHeight, remove_for_PlantHeight, remove_for_PlantArchitecture)
# remove the outliers # I can not remove the outliers
#plot(results_stu$StResiduals)
#outliers <- which(abs(scale(MyArchiphenos_replicated_filt$StResiduals))>4)
#MyArchiphenos_replicated_filtered <- MyArchiphenos_replicated_filt[-outliers,]

```
#My final pheno data is to bin the replicated and non_replicated
```{r}

```

## Run the model for each trait separately
#BranchlevelNum"
```{r}
MyArchiphenos_replicated <- MyArchiphenos %>% filter(Replicated == TRUE)
ArchiTRials <- MyArchiphenos_replicated %>% filter(!is.na(BranchlevelNum)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated  %>% 
       filter(!is.na(BranchlevelNum), BranchlevelNum !=0) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) != 0){
     try(MixedModels_BranchlevelNum <- lmer(BranchlevelNum ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_BranchlevelNum))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_BranchlevelNum))[,c("grp","vcov")] %>% .[2,2],
                      #VarRes = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels_BranchlevelNum))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_BranchlevelNum)/Mean, 
                      PEV = mean((se.ranef(MixedModels_BranchlevelNum)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "BranchlevelNum"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
 }
results_BranchlevelNum

```

#PlantArchitecture"
```{r}
MyArchiphenos_non_rep <- MyArchiphenos %>% filter(Replicated == FALSE) %>% filter(entryType == "check") %>% dplyr::select(-c(N_unique_accession, N_accession))
## Remove trials with only one check and with non replicate check.
To_remove <- MyArchiphenos_non_rep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% Toremove)

ArchiTRials <- MyArchiphenos_non_rep %>% filter(!is.na(PlantArchitecture)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_non_rep %>% 
       filter(!is.na(PlantArchitecture), PlantArchitecture !=0) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == 	"19introLatAmeumu")
     if(nrow(Data) != 0){
     try(MixedModels_PlantArchitecture <- lmer(PlantArchitecture ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_PlantArchitecture))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_PlantArchitecture))[,c("grp","vcov")] %>% .[2,2],
                      #VarRes = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels_PlantArchitecture))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_PlantArchitecture)/Mean, 
                      PEV = mean((se.ranef(MixedModels_PlantArchitecture)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "PlantArchitecture"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
 }

results_PlantArchitecture

#write.table(results2, file = "Result/MeansVarCompMixedModels.csv", quote = F,
 #           sep = ",", row.names = F)
```

#FirstBranchHeight
```{r}
MyArchiphenos_replicated <- MyArchiphenos %>% filter(Replicated == TRUE)

ArchiTRials <- MyArchiphenos_replicated %>% filter(!is.na(FirstBranchHeight), FirstBranchHeight !=0) %>% filter(!is.na(germplasmName)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names


j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated %>% 
       filter(!is.na(FirstBranchHeight), FirstBranchHeight !=0) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName ==	"19.GS.C4B.PYT.135.UB")
     if(nrow(Data) != 0){
     try(MixedModels_FirstBranchHeight <- lmer(FirstBranchHeight ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_FirstBranchHeight))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_FirstBranchHeight))[,c("grp","vcov")] %>% .[2,2],
                      #VarRes = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels_FirstBranchHeight))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_FirstBranchHeight)/Mean, 
                      PEV = mean((se.ranef(MixedModels_FirstBranchHeight)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "FirstBranchHeight"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
 }

results_FirstBranchHeight
      
#write.table(results2, file = "Result/MeansVarCompMixedModels.csv", quote = F,
 #           sep = ",", row.names = F)
```
#PlantHeight"
```{r}
MyArchiphenos_non_rep <- MyArchiphenos %>% filter(Replicated == FALSE) %>% filter(entryType == "check") %>% dplyr::select(-c(N_unique_accession, N_accession))
## Remove trials with only one check and with non replicate check.
To_remove <- MyArchiphenos_non_rep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% Toremove)

ArchiTRials <- MyArchiphenos_non_rep %>% filter(!is.na(PlantHeight)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_non_rep  %>% 
       filter(!is.na(PlantHeight), PlantHeight !=0) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) > 1){
     try(MixedModels_PlantHeight <- lmer(PlantHeight ~ blockInrep + (1| germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_PlantHeight))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_PlantHeight))[,c("grp","vcov")] %>% .[2,2],
                      #VarRes = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels_PlantHeight))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_PlantHeight)/Mean, 
                      PEV = mean((se.ranef(MixedModels_PlantHeight)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "PlantHeight"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
 }
results_PlantHeight
# Error : number of levels of each grouping factor must be < number of observations (problems: germplasmName) is because all the accessions have only one observation per block ("21_YellowFreshAyt31_Namulonge"). boundary (singular) fit: see help('isSingular') is for the very low genetic variation ("19C2bCETnonGSumu"). Ex #"14.GS.C1.EC.329.MK", 

#write.table(results_PlantHeight, file = "Result/MixedModelResults.csv", quote = F, sep =",",
        #    row.names = F)
```
#Data distribution visualization with replicated anf filtered trials trials and outliers removed
```{r}
MyArchiphenos_replicated_filtered <-  read.csv("data/MyArchiphenos_replicated_filtered.csv")
#Architraits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "PlantArchitecture")

MyArchiphenos_replicated_filtered$trialType <- factor(MyArchiphenos_replicated_filtered$trialType, 
                                      levels = c("CET", "PYT", "AYT", "UYT")) # Reordering group factor levels
MyArchiphenos_replicated_filtered$ArchiTRait <- factor(MyArchiphenos_replicated_filtered$ArchiTRait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture"))
#Why AM I still having a big outlier with plantheight so I added my hand removing
MyArchiphenos_replicated_filtered <- MyArchiphenos_replicated_filtered %>% filter(!ArchiTRait == "PlantArchitecture")
ggplot(data = MyArchiphenos_replicated_filtered,
                             aes(x=trialType,y=ArchiValue, fill = ArchiTRait)) +
  geom_boxplot(stat = "boxplot") +
  stat_summary(fun=mean, na.rm = TRUE) +
  labs(x= "Tria type", y= "Values") + theme_bw() +
    facet_grid(ArchiTRait~trialType, scales = "free")
```

