---
title: "fieldtrialfilter"
author: "Opamelas83"
date: "2023-02-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## check the number of each trait observations by trial. because most of the trials have less than the 3 traits observed. 
```{r}
library(tidyverse); library(data.table); library(lme4); library(here); library(magrittr); library(lubridate); library(MuMIn); library(arm)
MyArchiphenos <- read_csv("data/MyArchiphenotypes.csv", col_names = TRUE)
#The pipeline version of this analysis will use the TRUE/FALSE values of CompleteBlocks and IncompleteBlocks (Preliminary analysis of trial data ).
MyArchiphenos %>% 
     count(Replicated,non_Replicated,Location) %>% 
     spread(Location,n)
```
#Convert the data to “long format” . Remove missing values. “Nest” the data by Trait.
```{r}
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "DMC", "FRWT", "FTWT")
MyArchiphenos <- MyArchiphenos %>% 
      # Convert the data to "long format" 
      pivot_longer(cols = all_of(ArchiTRaits), 
                   names_to = "ArchiTRait", values_to = "ArchiValue") %>%
      # Remove missing values 
      filter(!is.na(ArchiValue)) %>% 
    filter(!is.na(replicate)) %>% 
      filter(!is.na(germplasmName))
# for counting initial trials and germplasm "PlantArchitecture", "DMC", "FRWT", "FTWT", "MCMDS"length(unique(MyArchiphenosQ$studyName))[1] 352 length(unique(MyArchiphenosQ$germplasmName))[1] 14416
```

##Fit models for replicated trials 
```{r}
MyArchiphenos_replicated <- MyArchiphenos %>% filter(Replicated == TRUE) #%>% filter(!ArchiTRaits %in% "PlantArchitecture")
To_remove <- MyArchiphenos_replicated  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_replicate=length(unique(replicate))) %>%
  filter(N_unique_accession == 1 | N_unique_replicate == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_replicated <- MyArchiphenos_replicated %>% dplyr::filter(!studyName %in% Toremove) %>% mutate(blockNumber = as.factor(blockNumber))#Took to remove outliers
ArchiTRials <- MyArchiphenos_replicated$studyName %>% table %>% .[order(.)] %>% names
 
i <- ArchiTRaits[1]
j <- ArchiTRials[1]
results <- tibble()
results_stu <- tibble()
for(i in ArchiTRaits){
 for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated %>% 
       filter(studyName == j, ArchiTRait == i)
        Data <- Data %>% #number of levels of each  grouping factor must be < number of observations (problems: germplasmName)
       mutate(PlantsPerPlot = ifelse(!all(is.na(PHar)), max(PHar, na.rm=T), NA),
              PropPHar = PHar / PlantsPerPlot,
              ArchiValue = ifelse(!is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = (ArchiValue / (PlantsPerPlot*PlotArea))*10,
                                  no = ifelse(is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = NA, no = ArchiValue)),
              PropPHar  = PropPHar) %>% 
          filter(!is.na(ArchiValue))# determine the root and the top yield in t/ha 
      if(nrow(Data) > 1){
      if(length(Data$germplasmName) > length(unique(Data$germplasmName))){
      if(unique(Data$ArchiTRait) %in% c("FRWT","FTWT")){
       Data <- Data %>% filter(!is.na(PropPHar))} else{# na in ProPHar does not causes reduce of the stu_resid outcome.
         Data <- Data}
     Formula1 <- "ArchiValue ~ repInTrial+ (1|germplasmName)"
     Formula <- ifelse(unique(Data$ArchiTRait) %in% c("FRWT", "FTWT") & length(unique(Data$PropPHar)) > 1, yes = paste0(Formula1," + PropPHar"), Formula1)
#Error : number of levels of each grouping factor must be < number of observations (problems: germplasmName), this happended only for one trial #which(!SSS %in% www)
     if(length(unique(Data$replicate)) > 1){# after filtering some trials have only   1 rep for FTWT
     try(MixedModels <- lmer(as.formula(Formula), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trait = i,
                      Trial = j,
                          VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2],
                      
                      Mean = as.data.frame(fixef(MixedModels))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),2),
                      CV = sigma(MixedModels)/Mean,
                      PEV = mean((se.ranef(MixedModels)$germplasmName[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),2)))
 ##Calculate Studentized Residuals
 try(stud_resids <- tibble(StResiduals = stats::rstudent(MixedModels),
                           studyName = Data$studyName,
                           germplasmName = Data$germplasmName, 
                           plotNumber = Data$plotNumber, 
                           ArchiTRait = Data$ArchiTRait,
                           ArchiValue = Data$ArchiValue))
 ##Bind all results
     try(results <- bind_rows(results, result))
 try(results_stu <- bind_rows(results_stu, stud_resids))
     rm(MixedModels); rm(result); rm(stud_resids)
     }
     }
    }

 }
 }
results
results_stu
write.table(results, file = "Result/MixedModel_replicated.csv", quote = T, sep =",", append = FALSE, row.names = F)
write.table(results_stu, file = "data/St_residual_replicated.csv", quote = T, sep =",", append = FALSE, row.names = F)
```
# Filter trials for identify trials with low heritability (H2 < 0.1) and/or low accuracy (Ac < 0.5) for all the traits, then, removes the outliers with residual value > 3.
```{r}
#add studentized residuals to orignal dataset
results_stu <- read.csv("data/St_residuall_replicated.csv")
MyArchiphenos_replicated_St <- left_join(MyArchiphenos_replicated, results_stu)
write.table(MyArchiphenos_replicated_St, file = "data/MyArchiphenos_replicated_St.csv", quote = T, sep =",", append = FALSE, row.names = F)
#Filter with H2 and Ac
results <- read.csv("Result/MixedModel_replicated.csv")
# Visualize the distribution 
plot(results$H2, results$Ac) # It is a challenge to chose the value of H2 and AC
Trials_replicated_to_keep <- results %>% filter((!is.na(H2) & H2 > 0.1) | (!is.na(Ac) & Ac >= 0.5)) #VarG/H2 > 1 for high replicate level (> accession number? but = 6,10,9 ...)
#unstead of removing the trails which are not interesting from my my data, I will filter then keep trials that have been have interesting genetically. I keep the trials by trait since a given trial can be in the "remove position" for 1 trait but be in the "stay position" for another trait. 
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "DMC", "FRWT", "FTWT")#
MyArchiphenos_replicated_filtered <- tibble()
for(i in ArchiTRaits){
  subdata <- Trials_replicated_to_keep %>% filter(Trait == i) 
  subdata <- subdata$Trial
  MyArchiphenos_replicated_filt <- MyArchiphenos_replicated_St %>% filter(ArchiTRait == i) 
  MyArchiphenos_replicated_filtered1 <- MyArchiphenos_replicated_filt %>% dplyr::filter(studyName %in% subdata)
  MyArchiphenos_replicated_filtered <- bind_rows(MyArchiphenos_replicated_filtered, MyArchiphenos_replicated_filtered1)
}

# another way to remove trials :  Make a list of trial to remove per trait
#for(i in ArchiTRaits){
#  subdata <- results %>% filter(Trait == i) %>% 
 # filter(H2 <= 0.1 | H2 >= 1 | Ac <= 0.4) %>% dplyr::select(Trial)
#   paste(ArchiTRaits[i] <- subdata)}

#Listfor_BranchlevelNum <- ArchiTRaits$BranchlevelNum
#remove_for_BranchlevelNum <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "BranchlevelNum") %>% dplyr::filter(!studyName %in% Listfor_BranchlevelNum)
#Listfor_FirstBranchHeight <- ArchiTRaits$FirstBranchHeight
#remove_for_FirstBranchHeight <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "FirstBranchHeight") %>% dplyr::filter(!studyName %in% Listfor_FirstBranchHeight)
#Listfor_PlantHeight <- ArchiTRaits$PlantHeight
#remove_for_PlantHeight <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "PlantHeight") %>% dplyr::filter(!studyName %in% Listfor_PlantHeight)
#Listfor_PlantArchitecture <- ArchiTRaits$PlantArchitecture
#remove_for_PlantArchitecture <- MyArchiphenos_replicated_St %>% dplyr::filter(ArchiTRait == "PlantArchitecture") %>% dplyr::filter(!studyName %in% Listfor_PlantArchitecture)

#MyArchiphenos_replicated_filt <- bind_rows(remove_for_BranchlevelNum, #remove_for_FirstBranchHeight, remove_for_PlantHeight, remove_for_PlantArchitecture)

# remove the outliers 
plot(MyArchiphenos_replicated_filtered$StResiduals)
outliers <- which(abs(MyArchiphenos_replicated_filtered$StResiduals) > 3)
MyArchiphenos_replicated_filtered_outl <- MyArchiphenos_replicated_filtered[-outliers,]
write.table(MyArchiphenos_replicated_filtered_outl, file = "data/MyArchiphenos_replicated_filtered_outl.csv", quote = F, append = FALSE, sep = " ",row.names = F)
#plot(MyArchiphenos_replicated_filtered_outl$StResiduals)
```

##Fit models for the non_replicated trials: The method is from "Welcome to the Introduction to Augmented Design Webinar" Dr. Jennifer Kling Presentation & Supplemental Files www.extension.org/pages/60430 Brought to you by: Plant Breeding and Genomics
```{r}
MyArchiphenos_nonrep <- MyArchiphenos %>% filter(Replicated == FALSE)
# add new entries random;set anyname 
MyArchiphenos_nonrep <- MyArchiphenos_nonrep %>% 
  mutate(New = ifelse(entryType == "check", 0, 1), 
         given_numb = row_number(),
         entryc = ifelse(entryType == "check", given_numb, 100000),
         entry = germplasmName,
         blockNumber = as.factor(blockNumber))
# Remove trials with only one accession or only 1 bloc=rep or trials without test/check
To_remove <- MyArchiphenos_nonrep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber)),
               N_unique_entryType=length(unique(entryType))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1 | N_unique_entryType == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_nonrep <- MyArchiphenos_nonrep %>% dplyr::filter(!studyName %in% Toremove)
# Some trials like "21_YellowFreshAyt31_Namulonge" do not have replicated checks
germplasmtoremove <- MyArchiphenos_nonrep  %>% 
  filter(entryType == "check") %>% 
  group_by(studyName, germplasmName) %>% 
  summarize(N_observation = length(germplasmName)) %>% 
  filter(N_observation == 1) #to check the trials including germplasm/check with 1 value
germplasmtoremove_trials <- germplasmtoremove$studyName
MyArchiphenos_nonrep <- MyArchiphenos_nonrep %>% dplyr::filter(!studyName %in% germplasmtoremove_trials)
##to check the trials including germplasm with 1 value: Is it necessary?
#germplasmtoremove_trials <- germplasmtoremove$studyName
#germplasmtoremove_germ <- germplasmtoremove$germplasmName
#MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% filter((!studyName %in% germplasmtoremove_trials) & (!germplasmName %in% germplasmtoremove_germ))
#MyArchiphenos_nonrep <- MyArchiphenos_nonrep %>% filter(ArchiTRait == "FirstBranchHeight") 
ArchiTRials <- MyArchiphenos_nonrep %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

i <- ArchiTRaits[1]
j <- ArchiTRials[1]
results_nonrep <- tibble()
results_stu_nonrep <- tibble()
for(i in ArchiTRaits){
 for(j in ArchiTRials){
     Data <- MyArchiphenos_nonrep %>% 
       filter(studyName == j , ArchiTRait == i)
       Data <- Data %>% 
       mutate(PlantsPerPlot = ifelse(!all(is.na(PHar)), max(PHar, na.rm=T), NA),
              PropPHar = PHar / PlantsPerPlot,
              ArchiValue = ifelse(!is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = (ArchiValue / (PlantsPerPlot*PlotArea))*10,
                                  no = ifelse(is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = NA, no = ArchiValue)),
              PropPHar  = PropPHar,
              New = as.numeric(New)) %>% 
          filter(!is.na(ArchiValue))
     if(nrow(Data) > 1){
       #Error : number of levels of each grouping factor must be < number of observations (problems: germplasmName), this happended only for one trial  
      if(length(Data$germplasmName) > length(unique(Data$germplasmName))){ 
      if(unique(Data$ArchiTRait) %in% c("FRWT","FTWT")){
       Data <- Data %>% filter(!is.na(PropPHar))} else {
         Data <- Data}
 Formula1 <- "ArchiValue ~ entryc + (1|blockNumber) + (1|entry:New)"
  Formula <- ifelse(unique(Data$ArchiTRait) %in% c("FRWT", "FTWT") & length(unique(Data$PropPHar)) > 1, yes = paste0(Formula1," + PropPHar"), Formula1)
# if(length(unique(Data$blockNumber)) > 1){
#if(nrow(Data <- Data %>% filter(New == 0)) > 1){#trials with check <  1
# if(nrow(Data <- Data %>% filter(New == 1)) > 1){ #trials with test  <  1
try(MixedModels <- lmer(as.formula(Formula), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trait = i,
                      Trial = j,
                          VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),1),
                      CV = sigma(MixedModels)/Mean,
                      PEV = mean((se.ranef(MixedModels)$entry[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),1)
 ))
##Calculate Studentized Residuals
try(stud_resids <- tibble(StResiduals = stats::rstudent(MixedModels),
                       studyName = Data$studyName,
                       germplasmName = Data$germplasmName,
                       plotNumber = Data$plotNumber,
                       ArchiTRait = Data$ArchiTRait,
                       ArchiValue = Data$ArchiValue))
 ##Bind all results
 try(results_nonrep <- bind_rows(results_nonrep, result))
 try(results_stu_nonrep <- bind_rows(results_stu_nonrep, stud_resids))
     (MixedModels); rm(result); rm(stud_resids)
      }
      }
   }
    }

results_nonrep
results_stu_nonrep
write.table(results_nonrep, file = "Result/MixedModel_nonreplicated.csv", quote = TRUE, sep =",", append = FALSE, row.names = F)

```
# Filter trials for identify trials with low heritability (H2 < 0.1) and/or low accuracy (Ac < 0.5) for all the traits
```{r}
#add studentized residuals to orignal dataset:
MyArchiphenos_nonrep_St <- left_join(MyArchiphenos_nonrep, results_stu_nonrep)
write.table(MyArchiphenos_nonrep_St, file = "data/St_residual_nonreplicated.csv", quote = TRUE, sep =",", append = FALSE, row.names = F)
#Filter with H2 and Ac
results_nonrep <- read_csv("Result/MixedModel_nonreplicated.csv", col_names = TRUE)
# Determine the limit values of H2 and Ac by trait 
results_nonrep %>% filter(Trait == "FirstBranchHeight") %>% summary()
# Visualize the distribution 
plot(results_nonrep$H2, results_nonrep$Ac) # It is a challenge to chose the value of H2 and AC. Here the trials have high values of both.
Trials_nonreplicated_to_keep <- results_nonrep %>% filter((!is.na(H2) & H2 > 0.1) | (!is.na(Ac) & Ac >= 0.5)) # I should remove the trials by trait since a given trial can be in the "remove position" for 1 trait but be in the "stay position" for another trait. 
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "DMC", "FRWT", "FTWT")#

MyArchiphenos_nonreplicated_filtered <- tibble()
for(i in ArchiTRaits){
  subdata <- Trials_nonreplicated_to_keep %>% filter(Trait == i) 
  subdata <- subdata$Trial
  MyArchiphenos_nonreplicated_filt <- MyArchiphenos_nonrep_St %>% filter(ArchiTRait == i) 
  MyArchiphenos_nonreplicated_filtered1 <- MyArchiphenos_nonreplicated_filt %>% dplyr::filter(studyName %in% subdata)
  MyArchiphenos_nonreplicated_filtered <- bind_rows(MyArchiphenos_nonreplicated_filtered, MyArchiphenos_nonreplicated_filtered1)
}

# remove the outliers 
plot(MyArchiphenos_nonreplicated_filtered$StResiduals)
outliers <- which(abs(MyArchiphenos_nonreplicated_filtered$StResiduals) > 3)
MyArchiphenos_nonreplicated_filtered_outl <- MyArchiphenos_nonreplicated_filtered[-outliers,]
write.table(MyArchiphenos_nonreplicated_filtered_outl, file = "data/MyArchiphenos_nonreplicated_filtered_outl.csv", quote = TRUE, sep =",", append = FALSE, row.names = F)

#plot(MyArchiphenos_nonreplicated_filtered_outl$StResiduals)
```
#My final pheno data is to bin the replicated and non_replicated
```{r}
#MyArchiphenos_replicated_filtered_outl <- MyArchiphenos_replicated_filtered_outl %>% mutate(blockNumber = as.factor(blockNumber))
MyArchiphenos_final <- bind_rows(MyArchiphenos_replicated_filtered_outl, MyArchiphenos_nonreplicated_filtered_outl) 
write.table(MyArchiphenos_final, file = "data/MyArchiphenos_final.csv", quote = TRUE, sep =",", append = FALSE, row.names = F)
```

####Run the model for Plant architecture replicated trials
```{r}
MyArchiphenos <- read_csv("data/MyArchiphenotypes.csv", col_names = TRUE)
MyArchiphenos_replicated_Ar <- unique(MyArchiphenos$ArchiTRait) %>% filter(Replicated == TRUE) %>% filter(ArchiTRait == "PlantArchitecture") %>% mutate(ArchiValue = as.factor(ArchiValue))

ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "DMC", "FRWT", "FTWT")
MyArchiphenos <- MyArchiphenos %>% 
     # Convert the data to "long format" 
     pivot_longer(cols = all_of(ArchiTRaits), 
                  names_to = "ArchiTRait", values_to = "ArchiValue") %>%
     # Remove missing values 
     filter(!is.na(ArchiValue)) %>% 
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName))

#MyArchiphenos_replicated <- MyArchiphenos_replicated %>% filter(!is.na(ArchiValue)) #%>% mutate(ArchiValue = ifelse(ArchiValue == 1, 0, 1))
ArchiTRials <- MyArchiphenos_replicated_Ar %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names
j <- ArchiTRials[50]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated_Ar  %>% 
      # filter(!is.na(ArchiValue)) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) > 1){
      if(length(unique(Data$ArchiValue)) > 1){ 
     try(MixedModels_Archit <- glmer(ArchiValue ~ repInTrial+  + (1|germplasmName), data = Data,  family = binomial, nAGQ = 2))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trial = j,
                          VarG = as.data.frame(VarCorr(MixedModels_Archit))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(var(residuals(MixedModels_Archit))) %>% .[1,1],
                      
                      Mean = as.data.frame(fixef(MixedModels_Archit))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),2),
                      CV = sigma(MixedModels_Archit)/Mean,
                      PEV = mean((se.ranef(MixedModels_Archit)$germplasmName[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),2)))
     try(results_Archi <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
      }
  }   
}

results_Archi
write.table(results_Archi, file = "Result/MixedModels_Archi.csv", quote = T, sep =",", append = FALSE, row.names = F)

```
## Run the model for each trait separately
#PlantArchitecture replicated trials
```{r}
MyArchiphenos <- read_csv("data/MyArchiphenotypes.csv", col_names = TRUE)
MyArchiphenos_replicated <- MyArchiphenos %>% filter(Replicated == TRUE)
ArchiTRials <- MyArchiphenos_replicated %>% filter(!is.na(PlantArchitecture)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated  %>% 
      # filter(!is.na(PlantArchitecture)) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) != 0){
     try(MixedModels_Archit <- lmer(PlantArchitecture ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trial = j,
                          VarG = as.data.frame(VarCorr(MixedModels_Archit))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_Archit))[,c("grp","vcov")] %>% .[2,2],
                      
                      Mean = as.data.frame(fixef(MixedModels_Archit))[1,],
                      H2 = round(VarG/(VarG + VarE/meanNreps),2),
                      CV = sigma(MixedModels_Archit)/Mean,
                      PEV = mean((se.ranef(MixedModels_Archit)$germplasmName[,1])^2),
                      Ac = round((1-PEV/VarG)^(1/2),2)))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
}

results
write.table(results, file = "Result/MixedModels_Archit.csv", quote = T, sep =",", append = FALSE, row.names = F)

```
#PlantArchitecture non-replicated trials
```{r}
MyArchiphenos_non_rep <- MyArchiphenos %>% filter(Replicated == FALSE) %>% filter(entryType == "check") %>% dplyr::select(-c(N_unique_accession, N_accession))
## Remove trials with only one check and with non replicate check.
To_remove <- MyArchiphenos_non_rep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1) %>% dplyr::select(studyName)
Toremove <- To_remove$studyName
MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% Toremove)

ArchiTRials <- MyArchiphenos_non_rep %>% filter(!is.na(PlantArchitecture)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_non_rep  %>% 
       filter(!is.na(ArchiValue), ArchiValue !=0) %>%
   filter(!is.na(blockNumber)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) != 0){
     try(MixedModels_PlantArchitecture <- lmer(ArchiValue ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_PlantArchitecture))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_PlantArchitecture))[,c("grp","vcov")] %>% .[2,2],
                      #VarRes = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels_PlantArchitecture))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_PlantArchitecture)/Mean, 
                      PEV = mean((se.ranef(MixedModels_PlantArchitecture)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "PlantArchitecture"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
 }

results_PlantArchitecture

#write.table(results2, file = "Result/MeansVarCompMixedModels.csv", quote = F,
 #           sep = ",", row.names = F)
```
#BranchlevelNum"
```{r}
MyArchiphenos_replicated <- MyArchiphenos %>% filter(Replicated == TRUE)
ArchiTRials <- MyArchiphenos_replicated %>% filter(!is.na(BranchlevelNum)) %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_replicated  %>% 
       filter(!is.na(BranchlevelNum), BranchlevelNum !=0) %>%
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) != 0){
     try(MixedModels_BranchlevelNum <- lmer(BranchlevelNum ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_BranchlevelNum))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_BranchlevelNum))[,c("grp","vcov")] %>% .[2,2],
                      #VarRes = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2],
                      Mean = as.data.frame(fixef(MixedModels_BranchlevelNum))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_BranchlevelNum)/Mean, 
                      PEV = mean((se.ranef(MixedModels_BranchlevelNum)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "BranchlevelNum"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     #rm(MixedModels); rm(result); rm(MixedModelsrd)
     }
 }
results_BranchlevelNum

```
#FirstBranchHeight
```{r}
MyArchiphenos_non_rep <- MyArchiphenos %>% filter(Replicated == FALSE) %>% filter(entryType == "check") %>% dplyr::select(-c(N_unique_accession, N_accession))
## Remove trials with only one check and with non replicate check.
To_remove <- MyArchiphenos_non_rep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1) %>% dplyr::select(studyName)

Toremove <- To_remove$studyName
MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% Toremove)

ArchiTRials <- MyArchiphenos_non_rep %>% filter(ArchiTRait == "FirstBranchHeight") %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_non_rep  %>% 
       filter(!is.na(ArchiValue), ArchiValue !=0) %>%
   filter(!is.na(blockNumber)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j)
     if(nrow(Data) != 0){
     try(MixedModels_FirstBranchHeight <- lmer(ArchiValue ~ blockInrep + (1|germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_FirstBranchHeight))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_FirstBranchHeight))[,c("grp","vcov")] %>% .[2,2],
                      Mean = as.data.frame(fixef(MixedModels_FirstBranchHeight))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_FirstBranchHeight)/Mean, 
                      PEV = mean((se.ranef(MixedModels_FirstBranchHeight)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "FirstBranchHeight"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     rm(MixedModels); rm(result)
     }
 }

results_FirstBranchHeight
      
#write.table(results2, file = "Result/MeansVarCompMixedModels.csv", quote = F,
 #           sep = ",", row.names = F)
```
#PlantHeight"
```{r}
MyArchiphenos_non_rep <- MyArchiphenos %>% filter(Replicated == FALSE) %>% filter(entryType == "check") %>% dplyr::select(-c(N_unique_accession, N_accession))
## Remove trials with only one check and with non replicate check.
To_remove <- MyArchiphenos_non_rep  %>% group_by(studyName) %>% 
     summarize(N_unique_accession=length(unique(germplasmName)),
               N_unique_bloc=length(unique(blockNumber))) %>%
  filter(N_unique_accession == 1 | N_unique_bloc == 1) %>% dplyr::select(studyName)
To_remove <- To_remove %>% add_row(studyName = c("21_YellowFreshAyt31_Namulonge","21Pyt78_Namulonge"))
Toremove <- To_remove$studyName
MyArchiphenos_non_rep <- MyArchiphenos_non_rep %>% dplyr::filter(!studyName %in% Toremove)

ArchiTRials <- MyArchiphenos_non_rep %>% filter(ArchiTRait == "PlantHeight") %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names

j <- ArchiTRials[1]
results <- tibble()
for(j in ArchiTRials){
     Data <- MyArchiphenos_non_rep  %>% 
       filter(!is.na(ArchiValue), ArchiValue !=0) %>%
   filter(!is.na(blockNumber)) %>% 
     filter(!is.na(germplasmName)) %>% 
       filter(studyName == j) #remove"21_YellowFreshAyt31_Namulonge" and "21Pyt78_Namulonge"
     if(nrow(Data) > 1){
     try(MixedModels_PlantHeight <- lmer(ArchiValue ~ blockInrep + (1| germplasmName), data = Data))
 meanNreps<- Data  %>%
      count(germplasmName) %$% mean(n)
 try(result <- tibble(Trials = j,
                          VarG = as.data.frame(VarCorr(MixedModels_PlantHeight))[,c("grp","vcov")] %>% .[1,2],
                          VarE = as.data.frame(VarCorr(MixedModels_PlantHeight))[,c("grp","vcov")] %>% .[2,2],
                      Mean = as.data.frame(fixef(MixedModels_PlantHeight))[1,],
                      H2 = VarG/(VarG + VarE/meanNreps),
                      CV = sigma(MixedModels_PlantHeight)/Mean, 
                      PEV = mean((se.ranef(MixedModels_PlantHeight)$germplasmName[,1])^2), 
                      Ac = (1-PEV/VarG)^(1/2),
                      Trait = "PlantHeight"
                      #Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModel_PlantHeight)[2]),Thus, Ac = accuracy of the individual prediction = Reliability^(1/2) of each BLUP value. 
 ))
     try(results <- bind_rows(results, result))
     rm(MixedModels); rm(result)
     }
 }
results_PlantHeight
# Error : number of levels of each grouping factor must be < number of observations (problems: germplasmName) is because all the accessions have only one observation per block ("21_YellowFreshAyt31_Namulonge"). boundary (singular) fit: see help('isSingular') is for the very low genetic variation ("19C2bCETnonGSumu"). Ex #"14.GS.C1.EC.329.MK", 
```
#Data distribution visualization with replicated anf filtered trials trials and outliers removed
```{r}
MyArchiphenos_replicated_filtered <-  read.csv("data/MyArchiphenos_replicated_filtered.csv")
#Architraits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "PlantArchitecture")

MyArchiphenos_replicated_filtered_outl$trialType <- factor(MyArchiphenos_replicated_filtered_outl$trialType, 
                                     levels = c("CET", "PYT", "AYT", "UYT")) # Reordering group factor levels
MyArchiphenos_replicated_filtered_outl$ArchiTRait <- factor(MyArchiphenos_replicated_filtered_outl$ArchiTRait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture"))
MyArchiphenos_replicated_filtered_outl <- MyArchiphenos_replicated_filtered_outl %>% filter(!ArchiTRait == "PlantArchitecture")
ggplot(data = MyArchiphenos_replicated_filtered_outl,
                             aes(x=trialType,y=ArchiValue, fill = ArchiTRait)) +
  geom_boxplot(stat = "boxplot") +
  stat_summary(fun=mean, na.rm = TRUE) +
  labs(x= "Tria type", y= "Values") + theme_bw() +
    facet_grid(ArchiTRait~trialType, scales = "free")
```
