---
title: "generalanalysis"
author: "Opamelas83"
date: "2023-04-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

##1. Analysis with results directly shown on the manuscrit
```{r}
library(tidyverse); library(lme4); library(data.table); library(here); library(magrittr); library(lubridate); library(MuMIn); library(arm); library(gt); library(reactablefmtr); library(corrplot); library(factoextra); library(FactoMineR); library(purrr); library(lemon)

MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE) 
#length(unique(MyArchiphenos_final$germplasmName)) 9062
#length(unique(MyArchiphenos_final$studyName)) 274
##Summary table for Archipheno trait
MyArchiphenos_finalW <- MyArchiphenos_final  %>%
  mutate(ArchiTRait = recode(ArchiTRait,
      "FRWT" = "FYLD", "FTWT" = "TYLD")) %>% 
  dplyr::select(programName, Year, ArchiTRait, 
                Location, studyDesign,    studyName, germplasmName) %>% 
  rename(`Breeding program` = programName,   
         Trait = ArchiTRait, `Trial design` = studyDesign)#%>% filter(!ArchiTRait %in% c("DMC", "FYLD", "TYLD"))
#length(unique(MyArchiphenos_finalW$germplasmName))[1] 7725 #Different from the previous because of plant shape trait
#length(unique(MyArchiphenos_finalW$studyName))[1] 239
#length(unique(MyArchiphenos_finalW$Location)) 27
MySummaryData  <- MyArchiphenos_finalW %>% group_by(Trait, `Breeding program`, `Trial design`) %>% summarise(Year= length(unique(Year)), Location = length(unique(Location)), `nb Trials` = length(unique(studyName)), `nb clones` = length(unique(germplasmName)))
write_csv(MySummaryData, "Result/MySummaryData.csv", append = FALSE)
#want the number of trials for DMC, FYLD, and FTWT by breeding program, to complete the manuscript, it is too long to add by Trial design.
MySummaryData_Yield <- MyArchiphenos_finalW %>%
  filter(Trait %in% c("DMC", "FYLD", "TYLD")) %>%
  group_by(Trait, `Breeding program`) %>%
  summarise(`Number of trials` = n_distinct(studyName),
    .groups = "drop")
MySummaryData_Yield
##Summary table for Shape trait
Shapephenos_filtered <- read_csv("data/Shapephenos_filtered.csv", col_names = TRUE) 
#length(unique(Shapephenos_filtered$studyName)) 18
#length(unique(Shapephenos_filtered$germplasmName)) 1932
#length(unique(Shapephenos_filtered$Location))6
#28 locations = Shapephenos_filteredLocation U Shapephenos_filtered$Location
#A$germplasmName <- Shapephenos_filtered$germplasmName
#B$germplasmName <- MyArchiphenos_finalW$germplasmName
#AB <- c(A,B) length(unique(AB)) 9214
Shapephenos_filteredS <- Shapephenos_filtered %>% dplyr::select(programName, Year, Location, studyDesign, studyName, germplasmName) %>% rename(`Breeding program` = programName, `Trial design` = studyDesign)
MySummaryShapeDataS  <- Shapephenos_filteredS %>% group_by(`Breeding program`, `Trial design`) %>% summarise(Year= length(unique(Year)), Location = length(unique(Location)), `nb Trials` = length(unique(studyName)), `nb clones` = length(unique(germplasmName)))
write_csv(MySummaryShapeDataS , "Result/MySummaryShapeDataS.csv", append = FALSE)
## second method for trial only
MyArchiphenos_final <- MyArchiphenos_final %>% filter(!ArchiTRait %in% c("DMC", "FRWT", "FTWT")) %>% dplyr::select(programName, Year, ArchiTRait, Location, studyDesign, studyName) %>% unique() -> SummaryData1 #N_unique_accession,
SummaryData <- table(SummaryData1$ArchiTRait, SummaryData1$programName, SummaryData1$studyDesign, SummaryData1$Location, SummaryData1$Year) %>% as.data.frame()#SummaryData1$N_unique_accession
colnames(SummaryData) <- c("Trait", "Breeding program", "Trial design", "Location", "Year", "nb Trials")#"nb clones"
SummaryData <- SummaryData[SummaryData$`nb Trials`!=0,]
MySummaryData <- SummaryData %>% group_by(Trait,`Breeding program`, `Trial design`) %>% summarise(Year= length(unique(Year)), Location = length(unique(Location)), `nb Trials` = sum(`nb Trials`, na.rm = T))#`nb clones`=sum(`nb clones`, na.rm = T)
```

```{r}
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE) 
MyArchiphenos_final %>% dplyr::select(programName, Year, ArchiTRait, studyName) %>% unique() -> SummaryData1
SummaryData2 <- table(SummaryData1$ArchiTRait, SummaryData1$programName, SummaryData1$Year) %>% as.data.frame()
colnames(SummaryData2) <- c("ArchiTRait", "programName", "Year", "nb_Trials")
SummaryData2 <- SummaryData2[SummaryData2$`nb_Trials`!=0,]
SummaryData3 <- SummaryData2 %>% reactable(groupBy = c("ArchiTRait", "programName", "Year"), columns = list(
  "nb_Trials" = colDef(aggregate = "sum")), defaultPageSize = 15)
#New for presentation
SummaryData_present <- SummaryData2 %>% 
  group_by(ArchiTRait) %>%
  summarise(programName = length(unique(programName)),
            Year = length(Year), 
            nb_Trials = length(nb_Trials))
#reactablefmtr::save_reactable_test(SummaryData2, "SummaryData.html")
#reactablefmtr::save_reactable_test(SummaryData2, "SummaryData.png")
```
#2.Data distribution visualization with replicated and non_replicated trials and outliers removed
```{r}
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE) 
MyArchiphenos_finalf <- MyArchiphenos_final %>% filter(ArchiTRait %in% c("PlantHeight", "FirstBranchHeight", "BranchlevelNum")) %>% rename(Trait = ArchiTRait) #"PlantArchitecture"
# Mean Comparison of Breeding Programs for "PlantHeight", "FirstBranchHeight", and "BranchLevelNum Traits"
## mean value by trait by group
Meangroup <- MyArchiphenos_finalf %>% group_by(Trait, programName, trialType) %>%
  get_summary_stats(ArchiValue, type = "mean_sd")
write.csv(Meangroup , "Result/Meangroup.csv", append = FALSE)

finalf_AOV <- MyArchiphenos_finalf %>% 
  pivot_wider(names_from = Trait, values_from =ArchiValue) 
##ANOVA
PlantHeight_aov <- aov(PlantHeight ~ programName*trialType, data = finalf_AOV)
PlantHeight_aov_sum <- data.frame(summary(PlantHeight_aov)[[1]])
write.table(PlantHeight_aov_sum , "Result/PlantHeight_aov_sum.csv", append = FALSE, col.names = NA, row.names = TRUE)
PlantHeight_aov_Tuk <- TukeyHSD(PlantHeight_aov)
Table1 <- data.frame(PlantHeight_aov_Tuk$`programName:trialType`)
#write.table(Table1 , "T1.csv", append = FALSE)
FirstBranchHeight_aov <- aov(FirstBranchHeight ~ programName*trialType, data = finalf_AOV)
FirstBranchHeight_aov_sum <- data.frame(summary(FirstBranchHeight_aov)[[1]])
write.table(FirstBranchHeight_aov_sum , "Result/FirstBranchHeight_aov_sum.csv", append = FALSE)
FirstBranchHeight_aov_Tuk <- TukeyHSD(FirstBranchHeight_aov)
Table2 <- data.frame(FirstBranchHeight_aov_Tuk$`programName:trialType`)
#write.table(Table2 , "T2.csv", append = FALSE)
BranchlevelNum_aov <- aov(BranchlevelNum ~ trialType, data = finalf_AOV)
BranchlevelNum_aov_sum <- data.frame(summary(BranchlevelNum_aov)[[1]])
write.table(BranchlevelNum_aov_sum , "Result/BranchlevelNum_aov_sum.csv", append = FALSE)
BranchlevelNum_aov_Tuk <- TukeyHSD(BranchlevelNum_aov)
Table3 <- data.frame(BranchlevelNum_aov_Tuk$`trialType`)
#write.table(Table3 , "T3.csv", append = FALSE)
#summary.aov(General_AOV)

##Data distribution plot visualization
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE) 
MyArchiphenos_finalf <- MyArchiphenos_final %>% filter(ArchiTRait %in% c("PlantHeight", "FirstBranchHeight", "BranchlevelNum")) %>% rename(Trait = ArchiTRait)
MyArchiphenos_finalf_IITA <- MyArchiphenos_finalf %>% filter(programName %in% "IITA")
MyArchiphenos_finalf_IITA$trialType <- factor(MyArchiphenos_finalf_IITA$trialType,                                     levels = c("CET", "PYT", "AYT", "UYT"))
# Reordering group factor levels
MyArchiphenos_finalf_IITA$Trait <- factor(MyArchiphenos_finalf_IITA$Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum"))#"PlantArchitecture"
#jpeg(filename = "Result/IITA_Traits_distribution.jpeg")
 CompFIG1 <- ggplot(data = MyArchiphenos_finalf_IITA,
                             aes(x=trialType,y=ArchiValue, fill = Trait)) +
  geom_boxplot(stat = "boxplot") +
  stat_summary(fun=mean, na.rm = TRUE) +
  labs(x= "Trial type", y= "Values") + theme_bw() +
    facet_grid(Trait~trialType, scales = "free") +
   ggtitle("IITA")
  #geom_text(data = MyArchiphenos_finalf,
  #          aes(label = ArchiValue, y = ArchiValue + 0.08)) +
  #theme(legend.position='none')
#dev.off()

MyArchiphenos_finalf_NRCRI <- MyArchiphenos_finalf %>% filter(programName %in% "NRCRI") %>%
  filter(!(Trait %in% "BranchlevelNum"))

  MyArchiphenos_finalf_NRCRI$trialType <- factor(MyArchiphenos_finalf_NRCRI$trialType,                                     levels = c("CET", "PYT", "AYT", "UYT"))
# Reordering group factor levels
MyArchiphenos_finalf_NRCRI$Trait <- factor(MyArchiphenos_finalf_NRCRI$Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight"))#"PlantArchitecture""BranchlevelNum"
#jpeg(filename = "Result/NRCRI_Traits_distribution.jpeg")
 CompFIG2 <- ggplot(data = MyArchiphenos_finalf_NRCRI,
                             aes(x=trialType,y=ArchiValue, fill = Trait)) +
  geom_boxplot(stat = "boxplot") +
  stat_summary(fun=mean, na.rm = TRUE) +
  labs(x= "Trial type", y= "Values") + theme_bw() +
    facet_grid(Trait~trialType, scales = "free") +
   ggtitle("NRCRI")
 nt <- theme(legend.position='none')
pdf(
  file = "Result/TraitDistribution.pdf",
  width = 8,
  height = 5, onefile = FALSE)
#or tiff(filename = "Result/TraitDistribution.tiff",
#  width = 8, height = 5), units = "in", res = 600, compression = "lzw")
grid_arrange_shared_legend(
  CompFIG1 + nt,
  CompFIG2 + nt,
  ncol = 2,
  nrow = 1
)
dev.off()

#Residuals Distribution by trait
results <- read_csv("Result/MixedModel_replicated.csv", col_names = TRUE)
results_nonrep <- read_csv("Result/MixedModel_nonreplicated.csv", col_names = TRUE)
Allresults <- bind_rows(results, results_nonrep) %>% dplyr::select(Trial, Trait, VarE) %>% rename(ArchiTRait = Trait)
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE)
MyArchiphenos_final <- MyArchiphenos_final %>% group_by(trialType, studyName, ArchiTRait) %>% summarise(studyName = unique(studyName)) %>% rename(Trial = studyName)
My_residuals <- left_join(MyArchiphenos_final, Allresults) 
My_residuals <- My_residuals %>% filter(ArchiTRait %in% c("PlantHeight", "FirstBranchHeight", "BranchlevelNum"))#"PlantArchitecture"

My_residuals$trialType <- factor(My_residuals$trialType, 
                                      levels = c("CET", "PYT", "AYT", "UYT")) 
My_residuals$Trait <- factor(My_residuals$ArchiTRait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum"))# "PlantArchitecture"
ggplot(data = My_residuals,
                             aes(x=trialType,y=VarE, fill = ArchiTRait)) +
  geom_boxplot(stat = "boxplot") +
  stat_summary(fun=mean, na.rm = TRUE) +
  labs(x= "Tria type", y= "Residuals value") + theme_bw() +
    facet_grid(Trait~trialType, scales = "free")
```
# Distribution of plant shape
```{r}
MyArchiphenos <- read_csv("data/MyArchiphenotypes.csv", col_names = TRUE)
ShapeTRaits <- c("Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type")
Shapephenos <- MyArchiphenos %>% 
     # Convert the data to "long format" 
     pivot_longer(cols = all_of(ShapeTRaits), 
                  names_to = "ShapeTRait", values_to = "ShapeValue") %>%
     # Remove missing values 
     filter(!is.na(ShapeValue)) %>% 
   filter(!is.na(replicate)) %>% 
     filter(!is.na(germplasmName))

Shapephenos_plant <- Shapephenos %>% dplyr::select(Year, trialType, ShapeTRait, ShapeValue) %>% filter (!is.na(ShapeValue)) %>% mutate(`Plant type` = ShapeValue)
#ShapeTRait", values_to = "ShapeValue"
Shapephenos_plantsum <- Shapephenos_plant %>% filter(!(`Plant type`) %in% c("Non_Cylindrical", "Non_Umbrella", "Non_Open","Non_Compact")) %>% group_by(trialType, `Plant type`) %>% summarise(numb = length(`Plant type`)) %>% mutate(`Value (%)` = round((numb/rowsum(numb, trialType)) * 100)) #"Non_Erect"

Shapephenos_plantsum$trialType <- factor(Shapephenos_plantsum$trialType, 
                                      levels = c("CET", "PYT", "AYT", "UYT"))
Shapephenos_plantsum$`plant shape` <- factor(Shapephenos_plantsum$`Plant type`,      
                        levels = c("Compact", "Open", "Umbrella", "Cylindrical"))
#jpeg(filename = "Result/PlantShapedistri.jpeg")
pdf(
  file = "Result/PlantShapedistri.pdf",
  width = 8,
  height = 7
)

ggplot(data = Shapephenos_plantsum, aes(x = trialType, y = `Value (%)`, 
                                group = `plant shape`)) +
  geom_point(aes(color = `plant shape`), size = 3) +
  geom_line(aes(color=`plant shape`), linetype = 81) +
  scale_color_brewer(palette = "Dark2") +
 ggtitle("Plant shape") +
  xlab("breeding stage") #+
ggsave("Result/PlantShapedistri.png",  scale = 1,)
#theme(axis.title.x=element_blank())
dev.off()
```
###3.Correlation between trait using the BLUPs VALU
```{r}
# For calling original result objects 
results <- readRDS("Result/GeneralMML_result.rds")
results_Shape <- readRDS("Result/results_Shape.rds")
General_results <- bind_rows(results,results_Shape)
# COMPUTE SELIND FROM BLUPs
MML_blups <- General_results %>%
    dplyr::select(Trait,BLUPs) %>%
    unnest(BLUPs) %>%
    dplyr::select(GermplasmName,Trait,BLUP) %>%
    spread(Trait,BLUP)
# Correlation between trait using the BLUPs VALUE*instead of drgBLUPsfor the manuscript after reviewer inquiry
allBLUP <- MML_blups %>% # dplyr::select(-GermplasmName) %>%
  dplyr::select(PlantHeight, FirstBranchHeight, BranchlevelNum, Cylindrical, Umbrella, Open, Compact, DMC, FRWT, FTWT) %>% #na.omit()
  mutate(FYLD = FRWT, TYLD = FTWT) %>%
  dplyr::select(-FRWT, -FTWT)
allBLUP_cor <- cor(allBLUP, use='pairwise')
# SAVE FIGURE
pdf(
  file = "Result/BLUP_Correlation.pdf",
  width = 8,
  height = 7)

Figplot2 <- corrplot::corrplot(allBLUP_cor, method = "number", sig.level = 0.001, type = "upper", insig = "blank")
mtext("BLUP", at=6.5, line=1, cex=1.5) 
dev.off()
```
#4.Variance component
```{r}
results <- readRDS("Result/GeneralMML_result.rds")
results_Shape <- readRDS("Result/results_Shape.rds")
General_results <- bind_rows(results,results_Shape)
Vartable <- General_results %>% 
  dplyr::select(Trait, H2, VarComps) %>%
  unnest(VarComps) %>%
  dplyr::select(Trait, H2, germplasmName, Residual) %>%
  mutate(Vg = round(germplasmName, digits = 2), 
         Ve = round(Residual, digits = 2)) %>% 
  dplyr::select(Trait, H2, Vg, Ve) 
write_csv(Vartable, "Result/Vartable.csv", append = FALSE)
```

#Check the correlation between trait with the original data
```{r}
#MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE) 
#MyArchiphenos_final$trialType <- factor(MyArchiphenos_final$trialType, 
#                                      levels = c("CET", "PYT", "AYT", "UYT"))
#MyArchiphenos_final_cor <- MyArchiphenos_final %>% #filter(ArchiTRait == "PlantArchitecture")
#  dplyr::group_by(trialType, ArchiTRait, germplasmName) %>%
#  summarise(Value = mean(ArchiValue, na.rm = T)) 
#MyArchiphenos_final_cor <- MyArchiphenos_final_cor %>% spread(ArchiTRait, Value) #%>% na.omit()
#Cor by trial type
#MyArchiphenos_cor1 <- MyArchiphenos_final_cor %>%
#  filter(trialType == "CET") %>% ungroup() %>% dplyr::select (BranchlevelNum, FirstBranchHeight, PlantHeight,  DMC, FRWT, FTWT) %>% na.omit()#PlantArchitecture,, MCMDS
#MyArchiphenos_cor1 <- cor(MyArchiphenos_cor1)
#corrplot::corrplot(MyArchiphenos_cor1, method = "number", type #= "upper", sig.level = 0.01, insig = "blank")
#mtext("CET", at=6, line=1, cex=2)
```

# 5. PREPARE THE 3,113-ACCESSION GWAS GENOTYPE PANEL
```{r}
# Load genotype dosage matrix
dosages <- readRDS("data/dosages.rds") 
# Load the accession IDs retained for GWAS 
GWAS_ID <- readRDS(here::here("output", "GWAS_accession_IDs.rds")) 
# Keep only the accessions included in the GWAS panel 
GWAS_ID <- GWAS_ID[GWAS_ID %in% rownames(dosages)] 
dosages_GWAS <- dosages[GWAS_ID, , drop = FALSE] 
# Check number of individuals dim(dosages_GWAS) 
dim(dosages_GWAS) # 3113 61158 
length(unique(rownames(dosages_GWAS))) # 3113 
```
# 6. Assign breeding-program information to GWAS accessions
```{r}
## Phenotype-derived file containing breeding-program information: To get the breeding program name for all the genotypes, I downloaded a phenotype file including all the trials.
Access <- read_csv(here::here("data/Access_IITA_NRCRI.csv"))#genotypes from NRCRI and IITA
# Retain relevant accession/program information
AccessN <- Access %>% dplyr::select(-plotNumber, -blockNumber)
write_csv(AccessN, "data/Access_IITA_NRCRI_new.csv", append = FALSE)
AccessN <- read_csv("data/Access_IITA_NRCRI_new.csv")
# Identify accessions occurring in IITA, NRCRI, or both programs
Access2 <- AccessN %>%
  distinct(germplasmName, programName) %>%
  group_by(germplasmName) %>%
  summarise(programName = if_else(
      n_distinct(programName) > 1, "IITA and NRCRI",
      first(programName)), .groups = "drop")
```
# 7. Add breeding-program information to the GWAS dosage matrix
```{r}
#convert my 3,113-row dosage matrix
Mydosages_GWAS <- as.data.frame(dosages_GWAS) %>%
  rownames_to_column("germplasmName")
# Add breeding-program information
Mydosages_Pro <- Mydosages_GWAS %>%
  left_join(Access2, by = "germplasmName")
# Verify that all 3,113 GWAS accessions were retained
nrow(Mydosages_Pro)#: 3113
# Check for missing breeding-program information
sum(is.na(Mydosages_Pro$programName))#: 0
# Number of GWAS accessions represented by each breeding program
Mydosages_Pro %>% count(programName)#   programName    n
#1           IITA 3025
#2 IITA and NRCRI   83
#3          NRCRI    5
```
# 8. Prepare genotype matrix for PCA
```{r}
#Remove metadata columns and retain only SNP dosage data
Mydosages_Pca <- Mydosages_Pro %>%
  dplyr::select(-programName) %>%
  column_to_rownames(var = "germplasmName") %>%
  as.matrix()
# Verify dimensions
dim(Mydosages_Pca)#: 3113 61158
Mypca <- prcomp(Mydosages_Pca, center = TRUE, scale. = FALSE)
# Save PCA
saveRDS(Mypca,
  here::here("output", "GWAS_pca_result.rds"))

```
# 9. FIGURE 6 – POPULATION STRUCTURE OF THE GWAS PANEL
```{r}
Mypca <- readRDS("output/GWAS_pca_result.rds")
# Calculate variance explained
pca_variance <- 100 * (Mypca$sdev^2 /
    sum(Mypca$sdev^2))
pca_variance[1:3]# PC1 = 7.280030; PC2 = 3.689614; PC3 = 2.944112
# Add breeding-program labels and sample sizes for Figure 6
Mydosages_Pro$`Breeding Program` <- factor(
  Mydosages_Pro$programName,
  levels = c("IITA", "IITA and NRCRI", "NRCRI"),
  labels = c(
    "IITA (n = 3,025)",
    "IITA and NRCRI (n = 83)",
    "NRCRI (n = 5)"))
# Plot individuals from PCA
pdf(
  file = "Result/PCA_indiv.pdf",
  width = 9,
  height = 7)
print(
  fviz_pca_ind(
    Mypca,
    geom = "point",
    col.ind = Mydosages_Pro$`Breeding Program`,
    palette = "Dark2",
    legend.title = "Breeding Program",
    alpha.ind = 0.85,
    pointsize = 1.8
  ) +
    labs(
      x = paste0("PC1 (", round(pca_variance[1], 1), "%)"),
      y = paste0("PC2 (", round(pca_variance[2], 1), "%)")))
dev.off()
```
# 10.SUPPLEMENTARY FIGURE: PCA SCREE PLOTPCA scree plot
```{r}

pdf(
  file = "Result/PCA_screeplot.pdf",
  width = 9,
  height = 7)
fviz_eig(Mypca,
  addlabels = TRUE,
  ncp = 10)
dev.off()
```
# 11. SNP contributions to PC1 and PC2
```{r}
#SNP contribution analysis can remain
Mypca <- readRDS(here::here("output", "GWAS_pca_result.rds"))
pca_result <- get_pca(Mypca, element = c("var", "ind"))
saveRDS(pca_result, file = here::here("Result", "pca_result.rds"))
pca_result <- readRDS("Result/pca_result.rds")
pca_var <- get_pca_var(Mypca)$contrib
pca_var_dim <- as_tibble(pca_var, rownames = NA)
pca_var_dim <- rownames_to_column(pca_var_dim) %>%
  rename(marker = rowname)
pca_var_dim12 <- pca_var_dim %>%
  dplyr::select(marker, Dim.1, Dim.2)
# Save SNP contribution results
saveRDS(pca_var_dim12, here::here("Result",
    "GWAS_pca_marker_contributions_PC1_PC2.rds"))
```
# 8. ADD CHROMOSOME AND PHYSICAL POSITION TO SNP CONTRIBUTIONS
```{r}
# From dosage matrix CHANGED because Mydosages_Pca is now the PCA matrix for the exact 3,113 GWAS accessions.
dosagecolnamelist <- colnames(Mydosages_Pca)
dosagecolname <- as.data.frame(dosagecolnamelist) %>%
  rename(marker = dosagecolnamelist)
dosagecolname <- as_tibble(dosagecolname)
dosagemarker <- dosagecolname %>%
  separate(marker,
    c("chromosome", "position", "v1", "v2"),
    sep = "_", remove = FALSE) %>%
  dplyr::select(-v1, -v2) %>%
  mutate(position = as.numeric(position)) %>%
  mutate(position = round(position/100000, digits=1))
Dosage_pca <- left_join(pca_var_dim12, dosagemarker)
write_csv(Dosage_pca, "output/Dosage_pca.csv", append = FALSE)
```
# 12. FIGURE 7 – SNP CONTRIBUTIONS TO PC1 AND PC2
```{r}
Dosage_pca <- read_csv("output/Dosage_pca.csv")
Dosage_pca <- Dosage_pca %>%
  mutate(PC1 = Dim.1, PC2 = Dim.2,
    position = position/10, .keep = "unused")
Dosage_pca$chromosome <- factor(
  Dosage_pca$chromosome,
  levels = as.character(1:18))
# PC1
CompFIG1 <- ggplot(Dosage_pca, aes(x = position, y = `PC1`)) +
  geom_point(color = Dosage_pca$chromosome) +
  facet_wrap(~chromosome, ncol = length(unique(Dosage_pca$chromosome))) +
  scale_color_brewer(palette = "Dark2")

# PC2
CompFIG2 <- ggplot(Dosage_pca, aes(x = position, y = `PC2`)) +
  geom_point(color = Dosage_pca$chromosome) +
  facet_wrap(~chromosome, ncol = length(unique(Dosage_pca$chromosome))) +
  scale_color_brewer(palette = "Dark2")

# Combine PC1 and PC2
# library(ggpubr)
Figure7 <- ggarrange(CompFIG1,CompFIG2,
  ncol = 1, nrow = 2)
pdf(
  file = "Result/Figure7.pdf",
  width = 12,
  height = 7)
print(Figure7)
dev.off()
```
# 13. PCA INDIVIDUAL CONTRIBUTIONS USED IN ORIGINAL WORKFLOW
```{r}
# For GWAS analysis
pca_ind <- get_pca_ind(Mypca)
saveRDS(pca_ind, file = here::here(
    "Result", "pca_ind.rds"))
pca_ind <- pca_ind$contrib
pca_ind_dim <- as_tibble(pca_ind, rownames = NA)
pca_ind_dim <- rownames_to_column(pca_ind_dim) %>%
  rename(germplasmName = rowname)
pca_ind_dim12 <- pca_ind_dim %>%
  dplyr::select(germplasmName, Dim.1, Dim.2)
write_csv(pca_ind_dim12,
          here::here("Result", "pca_ind_dim12.csv"), append = FALSE)
```


