---
title: "generalanalysis"
author: "Opamelas83"
date: "2023-04-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Analysis with results directly shown on the manuscrit
```{r}
library(tidyverse); library(data.table); library(here); library(magrittr); library(lubridate); library(MuMIn); library(arm); library(gt); library(reactablefmtr); library(corrplot); library(factoextra); library(FactoMineR)
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE) 
#length(unique(MyArchiphenos_final$germplasmName)) 9062
#length(unique(MyArchiphenos_final$studyName)) 274
##Summary table for Archipheno trait
MyArchiphenos_finalW <- MyArchiphenos_final  %>%
  mutate(ArchiTRait = recode(ArchiTRait,
      "FRWT" = "FYLD", "FTWT" = "TYLD")) %>% 
  dplyr::select(programName, Year, ArchiTRait, 
                Location, studyDesign,    studyName, germplasmName) %>% 
  rename(`Breeding program` = programName,   
         Trait = ArchiTRait, `Trial design` = studyDesign)#%>% filter(!ArchiTRait %in% c("DMC", "FYLD", "TYLD"))
#length(unique(MyArchiphenos_finalW$germplasmName))[1] 7725 #Different from the previous because of plant shape trait
#length(unique(MyArchiphenos_finalW$studyName))[1] 239
#length(unique(MyArchiphenos_finalW$Location)) 27
MySummaryData  <- MyArchiphenos_finalW %>% group_by(Trait, `Breeding program`, `Trial design`) %>% summarise(Year= length(unique(Year)), Location = length(unique(Location)), `nb Trials` = length(unique(studyName)), `nb clones` = length(unique(germplasmName)))
write_csv(MySummaryData, "Result/MySummaryData.csv", append = FALSE)
#want the number of trials for DMC, FYLD, and FTWT by breeding program, to complete the manuscript, it is too long to add by Trial design.
MySummaryData_Yield <- MyArchiphenos_finalW %>%
  filter(Trait %in% c("DMC", "FYLD", "TYLD")) %>%
  group_by(Trait, `Breeding program`) %>%
  summarise(`Number of trials` = n_distinct(studyName),
    .groups = "drop")
MySummaryData_Yield
##Summary table for Shape trait
Shapephenos_filtered <- read_csv("data/Shapephenos_filtered.csv", col_names = TRUE) 
#length(unique(Shapephenos_filtered$studyName)) 18
#length(unique(Shapephenos_filtered$germplasmName)) 1932
#length(unique(Shapephenos_filtered$Location))6
#28 locations = Shapephenos_filteredLocation U Shapephenos_filtered$Location
#A$germplasmName <- Shapephenos_filtered$germplasmName
#B$germplasmName <- MyArchiphenos_finalW$germplasmName
#AB <- c(A,B) length(unique(AB)) 9214
Shapephenos_filteredS <- Shapephenos_filtered %>% dplyr::select(programName, Year, Location, studyDesign, studyName, germplasmName) %>% rename(`Breeding program` = programName, `Trial design` = studyDesign)
MySummaryShapeDataS  <- Shapephenos_filteredS %>% group_by(`Breeding program`, `Trial design`) %>% summarise(Year= length(unique(Year)), Location = length(unique(Location)), `nb Trials` = length(unique(studyName)), `nb clones` = length(unique(germplasmName)))
write_csv(MySummaryShapeDataS , "Result/MySummaryShapeDataS.csv", append = FALSE)
## second method for trial only
MyArchiphenos_final <- MyArchiphenos_final %>% filter(!ArchiTRait %in% c("DMC", "FRWT", "FTWT")) %>% dplyr::select(programName, Year, ArchiTRait, Location, studyDesign, studyName) %>% unique() -> SummaryData1 #N_unique_accession,
SummaryData <- table(SummaryData1$ArchiTRait, SummaryData1$programName, SummaryData1$studyDesign, SummaryData1$Location, SummaryData1$Year) %>% as.data.frame()#SummaryData1$N_unique_accession
colnames(SummaryData) <- c("Trait", "Breeding program", "Trial design", "Location", "Year", "nb Trials")#"nb clones"
SummaryData <- SummaryData[SummaryData$`nb Trials`!=0,]
MySummaryData <- SummaryData %>% group_by(Trait,`Breeding program`, `Trial design`) %>% summarise(Year= length(unique(Year)), Location = length(unique(Location)), `nb Trials` = sum(`nb Trials`, na.rm = T))#`nb clones`=sum(`nb clones`, na.rm = T)
```
# 1. PREPARE THE 3,113-ACCESSION GWAS GENOTYPE PANEL
```{r}
# Load genotype dosage matrix
dosages <- readRDS("data/dosages.rds") 
# Load the accession IDs retained for GWAS 
GWAS_ID <- readRDS(here::here("output", "GWAS_accession_IDs.rds")) 
# Keep only the accessions included in the GWAS panel 
GWAS_ID <- GWAS_ID[GWAS_ID %in% rownames(dosages)] 
dosages_GWAS <- dosages[GWAS_ID, , drop = FALSE] 
# Check number of individuals dim(dosages_GWAS) 
dim(dosages_GWAS) # 3113 61158 
length(unique(rownames(dosages_GWAS))) # 3113 
```
# 2. Assign breeding-program information to GWAS accessions
```{r}
## Phenotype-derived file containing breeding-program information: To get the breeding program name for all the genotypes, I downloaded a phenotype file including all the trials.
Access <- read_csv(here::here("data/Access_IITA_NRCRI.csv"))#genotypes from NRCRI and IITA
# Retain relevant accession/program information
AccessN <- Access %>% dplyr::select(-plotNumber, -blockNumber)
write_csv(AccessN, "data/Access_IITA_NRCRI_new.csv", append = FALSE)
AccessN <- read_csv("data/Access_IITA_NRCRI_new.csv")
# Identify accessions occurring in IITA, NRCRI, or both programs
Access2 <- AccessN %>%
  distinct(germplasmName, programName) %>%
  group_by(germplasmName) %>%
  summarise(programName = if_else(
      n_distinct(programName) > 1, "IITA and NRCRI",
      first(programName)), .groups = "drop")
```
# 3. Add breeding-program information to the GWAS dosage matrix
```{r}
#convert my 3,113-row dosage matrix
Mydosages_GWAS <- as.data.frame(dosages_GWAS) %>%
  rownames_to_column("germplasmName")
# Add breeding-program information
Mydosages_Pro <- Mydosages_GWAS %>%
  left_join(Access2, by = "germplasmName")
# Verify that all 3,113 GWAS accessions were retained
nrow(Mydosages_Pro)#: 3113
# Check for missing breeding-program information
sum(is.na(Mydosages_Pro$programName))#: 0
# Number of GWAS accessions represented by each breeding program
Mydosages_Pro %>% count(programName)#   programName    n
#1           IITA 3025
#2 IITA and NRCRI   83
#3          NRCRI    5
```
# 4. Prepare genotype matrix for PCA
```{r}
#Remove metadata columns and retain only SNP dosage data
Mydosages_Pca <- Mydosages_Pro %>%
  dplyr::select(-programName) %>%
  column_to_rownames(var = "germplasmName") %>%
  as.matrix()
# Verify dimensions
dim(Mydosages_Pca)#: 3113 61158
Mypca <- prcomp(Mydosages_Pca, center = TRUE, scale. = FALSE)
# Save PCA
saveRDS(Mypca,
  here::here("output", "GWAS_pca_result.rds"))
# Calculate variance explained
pca_variance <- 100 * (Mypca$sdev^2 /
    sum(Mypca$sdev^2))
pca_variance[1:3]# PC1 = 7.280030; PC2 = 3.689614; PC3 = 2.944112
```
# 5. FIGURE 6 – POPULATION STRUCTURE OF THE GWAS PANEL
```{r}
# Add breeding-program labels and sample sizes for Figure 6
Mydosages_Pro$`Breeding Program` <- factor(
  Mydosages_Pro$programName,
  levels = c("IITA", "IITA and NRCRI", "NRCRI"),
  labels = c(
    "IITA (n = 3,025)",
    "IITA and NRCRI (n = 83)",
    "NRCRI (n = 5)"))
# Plot individuals from PCA
jpeg(
  filename = "output/PCA_indiv.jpeg",
  width = 1800,
  height = 1400,
  res = 200)
print(
  fviz_pca_ind(
    Mypca,
    geom = "point",
    col.ind = Mydosages_Pro$`Breeding Program`,
    palette = "Dark2",
    legend.title = "Breeding Program",
    alpha.ind = 0.85,
    pointsize = 1.8
  ) +
    labs(
      x = paste0("PC1 (", round(pca_variance[1], 1), "%)"),
      y = paste0("PC2 (", round(pca_variance[2], 1), "%)")))
dev.off()

```
# 6.SUPPLEMENTARY FIGURE: PCA SCREE PLOTPCA scree plot
```{r}
jpeg(
  filename = "output/PCA_screeplot.jpeg",
  width = 1800,
  height = 1400,
  res = 200)
fviz_eig(Mypca,
  addlabels = TRUE,
  ncp = 10)
dev.off()
```
# 7. SNP contributions to PC1 and PC2
```{r}
#SNP contribution analysis can remain
Mypca <- readRDS(here::here("output", "GWAS_pca_result.rds"))
pca_result <- get_pca(Mypca, element = c("var", "ind"))
saveRDS(pca_result, file = here::here("Result", "pca_result.rds"))
pca_result <- readRDS("Result/pca_result.rds")
pca_var <- get_pca_var(Mypca)$contrib
pca_var_dim <- as_tibble(pca_var, rownames = NA)
pca_var_dim <- rownames_to_column(pca_var_dim) %>%
  rename(marker = rowname)
pca_var_dim12 <- pca_var_dim %>%
  dplyr::select(marker, Dim.1, Dim.2)
# Save SNP contribution results
saveRDS(pca_var_dim12, here::here("Result",
    "GWAS_pca_marker_contributions_PC1_PC2.rds"))
```
# 8. ADD CHROMOSOME AND PHYSICAL POSITION TO SNP CONTRIBUTIONS
```{r}
# From dosage matrix CHANGED because Mydosages_Pca is now the PCA matrix for the exact 3,113 GWAS accessions.
dosagecolnamelist <- colnames(Mydosages_Pca)
dosagecolname <- as.data.frame(dosagecolnamelist) %>%
  rename(marker = dosagecolnamelist)
dosagecolname <- as_tibble(dosagecolname)
dosagemarker <- dosagecolname %>%
  separate(marker,
    c("chromosome", "position", "v1", "v2"),
    sep = "_", remove = FALSE) %>%
  dplyr::select(-v1, -v2) %>%
  mutate(position = as.numeric(position)) %>%
  mutate(position = round(position/100000, digits=1))
Dosage_pca <- left_join(pca_var_dim12, dosagemarker)
write_csv(Dosage_pca, "output/Dosage_pca.csv", append = FALSE)
Dosage_pca <- read_csv("output/Dosage_pca.csv")
Dosage_pca <- Dosage_pca %>%
  mutate(PC1 = Dim.1, PC2 = Dim.2,
    position = position/10, .keep = "unused")
```
# 9. FIGURE 7 – SNP CONTRIBUTIONS TO PC1 AND PC2
```{r}
Dosage_pca$chromosome <- factor(
  Dosage_pca$chromosome,
  levels = as.character(1:18))
# PC1
CompFIG1 <- ggplot(Dosage_pca, aes(x = position, y = `PC1`)) +
  geom_point(color = Dosage_pca$chromosome) +
  facet_wrap(~chromosome, ncol = length(unique(Dosage_pca$chromosome))) +
  scale_color_brewer(palette = "Dark2")

# PC2
CompFIG2 <- ggplot(Dosage_pca, aes(x = position, y = `PC2`)) +
  geom_point(color = Dosage_pca$chromosome) +
  facet_wrap(~chromosome, ncol = length(unique(Dosage_pca$chromosome))) +
  scale_color_brewer(palette = "Dark2")

# Combine PC1 and PC2
# library(ggpubr)
Figure7 <- ggarrange(CompFIG1,CompFIG2,
  ncol = 1, nrow = 2)
Figure7
```
# 10. PCA INDIVIDUAL CONTRIBUTIONS USED IN ORIGINAL WORKFLOW
```{r}
# For GWAS analysis
pca_ind <- get_pca_ind(Mypca)
saveRDS(pca_ind, file = here::here(
    "Result", "pca_ind.rds"))
pca_ind <- pca_ind$contrib
pca_ind_dim <- as_tibble(pca_ind, rownames = NA)
pca_ind_dim <- rownames_to_column(pca_ind_dim) %>%
  rename(germplasmName = rowname)
pca_ind_dim12 <- pca_ind_dim %>%
  dplyr::select(germplasmName, Dim.1, Dim.2)
write_csv(pca_ind_dim12,
          here::here("Result", "pca_ind_dim12.csv"), append = FALSE)
```


