---
title: "genselect"
author: "Opamelas83"
date: "2023-04-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r}
install.packages("sommer") #new version is important (why?)
library(tidyverse); library(data.table); library(genomicMateSelectR); library(magrittr); library(sommer)
```

##K-fold Cross validation
```{r}
blups<-readRDS(here::here("output","blups.rds"))
dosages<-readRDS(file=here::here("data","dosages.rds"))
A<-readRDS(file=here::here("output","kinship_add.rds"))
D<-readRDS(file=here::here("output","kinship_AD.rds"))
D<-readRDS(file=here::here("output","kinship_dom.rds"))
grms[["D"]]<-D

blups %<>% 
     # need to rename the "blups" list to comply with the runCrossVal function
     rename(TrainingData=BLUPs) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     mutate(TrainingData=map(TrainingData,
                             ~filter(.,GermplasmName %in% rownames(A)) %>% 
                                  # rename the germplasmName column to GID
                                  rename(GID=GermplasmName)))

blups
## A tibble: 4 × 2
#  Trait             TrainingData        
#  <chr>             <list>              
#1 BranchlevelNum    <tibble [3,113 × 6]>
#2 FirstBranchHeight <tibble [2,698 × 6]>
#3 PlantHeight       <tibble [3,664 × 6]>
#4 PlantArchitecture <tibble [3,470 × 6]>
#5 Cylindrical_type  <tibble [1,518 × 6]>
#6 Umbrella_type     <tibble [1,045 × 6]>
#7 Open_type         <tibble [1,164 × 6]>
#8 Compact_type      <tibble [1,518 × 6]>
#9 Erectibility      <tibble [1,045 × 6]>


#For fastest, lightest compute of accuracy, remove non-phenotyped from kinship
#gids<-blups %>% 
#     unnest(TrainingData) %$% unique(GID)
# dim(A) [1] 963 963

#D<-D[gids,gids]

```
# selection indices:To think about 
```{r}
#Include selection index weights
SIwts<-c(PlantHeight=4,
         FirstBranchHeight=4,
         BranchlevelNum=4,
         PlantArchitecture=4)
#SIwts
```

#Execute cross-validation
```{r}
#For A 
standardCV<-runCrossVal(blups=blups,
                        modelType="A",
                        selInd=FALSE,
                        grms=list(A=A),
                        nrepeats=3,nfolds=5,
                        gid="GID",seed=999999999,
                        ncores=10)
#For "AD"
standardCV<-runCrossVal(blups=blups,
                        modelType="AD",
                        selInd=FALSE,
                        grms=list(A=A,D=D),
                        nrepeats=3,nfolds=5,
                        gid="GID",seed=999999999,
                        ncores=10)

#For "DirDim"
standardCV<-runCrossVal(blups=blups,
                        modelType="DirDom",
                        selInd=FALSE,
                        grms=list(A=A,D=D),
                        dosages=dosages,
                        nrepeats=3,nfolds=5,
                        gid="GID",seed=999999999,
                        ncores=8)
#standardCV %>% unnest(accuracyEstOut)
#Plot result
saveRDS(standardCV,file = here::here("output","AD_KfoldsCval.rds"))
jpeg(filename = "output/AD_KfoldCval.jpeg")

standardCV %>% 
     unnest(accuracyEstOut) %>% 
     mutate(Trait = factor(Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture", "Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type", "Erectibility")))%>%
     dplyr::select(repeats,id,predOf,Trait,Accuracy) %>% 
     ggplot(.,aes(x=Trait,y=Accuracy,fill=Trait)) + 
  ggtitle("AD_Kfold Cross validation") +
  geom_boxplot() + theme_bw()
dev.off()
# cross validation by breeding program
# cross validation by periode 

```

##Predict parental breeding values
```{r}
blups<-readRDS(here::here("output","blups.rds"))
A<-readRDS(file=here::here("output","kinship_add.rds"))

blups %<>% 
     # need to rename the "blups" list to comply with the runCrossVal function
     rename(TrainingData=BLUPs) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     mutate(TrainingData=map(TrainingData,
                             ~filter(.,GermplasmName %in% rownames(A)) %>% 
                                  # rename the germplasmName column to GID
                                  rename(GID=GermplasmName)))

blups
## A tibble: 4 × 2

```

```{r}
#Include selection index weights
SIwts<-c(PlantHeight=4,
         FirstBranchHeight=4,
         BranchlevelNum=4,
         PlantArchitecture=4)
#SIwts
```

Run genomic predictions
```{r}
gpreds<-runGenomicPredictions(modelType="A",
                              selInd=TRUE, SIwts=SIwts,
                              blups=blups,
                              grms=list(A=A),
                              ncores=3)
#Extract GEBV
gpreds
#We have a single-row tibble. To access a simple table listing GEBV for each trait and the selection index:
gpreds$gblups[[1]]

#At this point, you can use the SELIND predictions directly to rank and select parents. Example: sort by SELIND and pick the top 10?s[[1]]

gpreds$gblups[[1]] %>% 
     arrange(desc(SELIND)) %>% 
     slice(1:10)
#For more detailed output, including variance component estimates:

gpreds$genomicPredOut[[1]]

gpreds$genomicPredOut[[1]]$varcomps[[1]]
#>                         VarComp VarCompSE   Zratio

saveRDS(gpreds,file = here::here("output","genomicPredictions.rds"))
```

#Accuracy of cross prediction
```{r}
ped<-read_delim(here::here("data","archi_pedigree.txt"),delim = "\t")
# Keep only complete pedigree records
ped %<>% 
     dplyr::select(-Cross_Type) %>% 
     filter(!is.na(Female_Parent),
            !is.na(Male_Parent),
            Female_Parent!="?",
            Male_Parent!="?") %>% 
     distinct
# Number of full-sib families
ped %>% distinct(Female_Parent,Male_Parent) %>% nrow()
#1985
#Summarise distribution of full-sib family sizes
ped %>%   
  count(Female_Parent,Male_Parent) %>% arrange(desc(n))  %>%  summary(.$n)# We need 2 members by family for this analysis (why?)
#Full genotyped trio
dosages<-readRDS(here::here("data","dosages.rds"))
genotyped_gids<-rownames(dosages)
#Are all pedigree genotyped?
table(ped$Accession %in% genotyped_gids)
#FALSE  TRUE 
#  718  7982 
#FALSE
table(ped$Female_Parent %in% genotyped_gids)
#FALSE  TRUE 
# 1113  7587
#We need only parents and descendance present in our SNP data
genotyped_ped<-ped %>% 
     filter(Accession %in% genotyped_gids,
            Female_Parent %in% genotyped_gids,
            Male_Parent %in% genotyped_gids)
genotyped_ped %>% nrow()
# 6512 complete trios (accessions + male parent + female parents)
genotyped_ped %>% 
     count(Female_Parent,Male_Parent) %>% arrange(desc(n))  %>%  summary(.$n)
#Female_Parent      Male_Parent   looks like 1196 full-sib families.        # Length:1196        Length:1196  
genotyped_ped %>% 
     count(Female_Parent,Male_Parent) %>% 
     filter(n>1)#810 families have offsprinf >1.

```
# validate pedigree There is one additional step I highly recommend and will demonstrate here.Plant breeding pedigrees can often have errors, esp. for the male (pollen) parent. For that reason, I recommend using the genomic data to check the pedigree. We do not want our estimate of family-genetic variance prediction accuracy further detrimated by the presence of incorrect pedigree entrees.There are various software options to do this, probably an R package or two. My approach uses the --genome IBD calculator in the command-line program PLINK v1.9, click here for the PLINK1.9 manual and to download/install the program.
```{r}
#plink --vcf VCF/BreedBaseGenotypes.vcf.gz \
 #    --make-bed --const-fid --keep-allele-order \
  #   --out VCF/BreedBaseGenotypes_subset

#plink --bfile VCF/BreedBaseGenotypes_subset \
#  --genome \
 # --out ../output/BreedBaseGenotypes_subset;

```

```{r}
#verify parent-offspring
genome %>% headgenome<-read.table(here::here("output/","BreedBaseGenotypes_subset.genome"),
                   stringsAsFactors = F,header = T) %>% 
     as_tibble
genome %>% head
dim(genome)
# 49297485       14
ped %>% 
     semi_join(genome %>% rename(Accession=IID1,Female_Parent=IID2)) %>% 
     left_join(genome %>% rename(Accession=IID1,Female_Parent=IID2))
```

```{r}
#Confirm Female_Parent - Offspring Relationship
## In the plink genome file
## IID1 or IID2 could be the Accession or the Female_Parent
conf_female_ped<-genotyped_ped %>% 
     inner_join(genome %>% 
                     rename(Accession=IID1,Female_Parent=IID2)) %>% 
     bind_rows(genotyped_ped %>% 
                    inner_join(genome %>% 
                                    rename(Accession=IID2,Female_Parent=IID1))) %>% 
     # Declare confirm-reject Accession-Female_Parent
     mutate(ConfirmFemaleParent=case_when(Z0<0.32 & Z1>0.67~"Confirm", 
                                          # Relatedness coeff differ if the Accession is the result of a self-cross
                                          Male_Parent==Female_Parent & PI_HAT>0.6 & Z0<0.3 & Z2>0.32~"Confirm",
                                          TRUE~"Reject")) %>% 
     dplyr::select(Accession,Female_Parent,ConfirmFemaleParent)
#> Joining, by = c("Accession", "Female_Parent")
#> Joining, by = c("Accession", "Female_Parent")
## Now do the same for the Accession-Male_Parent relationships
conf_male_ped<-genotyped_ped %>% 
     inner_join(genome %>% 
                     rename(Accession=IID1,Male_Parent=IID2)) %>% 
     bind_rows(genotyped_ped %>% 
                    inner_join(genome %>% 
                                    rename(Accession=IID2,Male_Parent=IID1))) %>% 
     # Declare confirm-reject Accession-Female_Parent
     mutate(ConfirmMaleParent=case_when(Z0<0.32 & Z1>0.67~"Confirm", 
                                          # Relatedness coeff differ if the Accession is the result of a self-cross
                                          Male_Parent==Female_Parent & PI_HAT>0.6 & Z0<0.3 & Z2>0.32~"Confirm",
                                          TRUE~"Reject")) %>% 
     dplyr::select(Accession,Male_Parent,ConfirmMaleParent)
#> Joining, by = c("Accession", "Male_Parent")
#> Joining, by = c("Accession", "Male_Parent")
# Now join the confirmed female and male relationships
# This regenerates the original "genotyped_ped" with two added columns
conf_male_ped<-genotyped_ped %>% 
     inner_join(genome %>% 
                     rename(Accession=IID1,Male_Parent=IID2)) %>% 
     bind_rows(genotyped_ped %>% 
                    inner_join(genome %>% 
                                    rename(Accession=IID2,Male_Parent=IID1))) %>% 
     # Declare confirm-reject Accession-Female_Parent
     mutate(ConfirmMaleParent=case_when(Z0<0.32 & Z1>0.67~"Confirm", 
                                          # Relatedness coeff differ if the Accession is the result of a self-cross
                                          Male_Parent==Female_Parent & PI_HAT>0.6 & Z0<0.3 & Z2>0.32~"Confirm",
                                          TRUE~"Reject")) %>% 
     dplyr::select(Accession,Male_Parent,ConfirmMaleParent)
#> Joining, by = c("Accession", "Male_Parent")
#> Joining, by = c("Accession", "Male_Parent")
# Now join the confirmed female and male relationships
# This regenerates the original "genotyped_ped" with two added columns
confirmed_ped<-conf_female_ped %>% 
     left_join(conf_male_ped) %>% 
     relocate(Male_Parent,.before = "ConfirmFemaleParent")
#> Joining, by = "Accession"
```

```{r}
#how well supported are the pedigree relationships according to my approach?
confirmed_ped %>% 
     count(ConfirmFemaleParent,ConfirmMaleParent) %>% 
  mutate(Prop=round(n/sum(n),2))
#  A tibble: 4 × 4
#  ConfirmFemaleParent ConfirmMaleParent     n  Prop
#  <chr>               <chr>             <int> <dbl>
#1 Confirm             Confirm            4503  0.69
#2 Confirm             Reject              888  0.14
#3 Reject              Confirm             577  0.09
#4 Reject              Reject              544  0.08    
```
Subset to fully-validated trios
We can only run the cross-validation using a pedigree where the full trio (Accession?s relationship to both parents) is validated.
Remove any without both parents confirmed.
```{r}
valid_ped<-confirmed_ped %>% 
     filter(ConfirmFemaleParent=="Confirm",
         ConfirmMaleParent=="Confirm") %>% 
     dplyr::select(-contains("Confirm"))
valid_ped %>% nrow()# 4503

valid_ped %>% 
    count(Female_Parent,Male_Parent) %>% 
    filter(n>1)
# 565 × 3; 565 full-sib families that have >1 entry are confirmed.

valid_ped %>% 
    count(Female_Parent,Male_Parent) %>% 
    filter(n>2)
# 437 × 3; 437 families have more than 2 entry
saveRDS(valid_ped,here::here("output","verified_ped.rds"))
```
Parent-wise cross-validation
Refer to:
genomicMateSelectR::runParentWiseCrossVal() documentation
Example of IITA_2021GS Cross-validationParent-wise cross-validation
```{r}
ped<-readRDS(here::here("output","verified_ped.rds")) %>% 
     # Rename things to match genomicMateSelectR::runParentWiseCrossVal()
     rename(GID=Accession,
            sireID=Male_Parent,
            damID=Female_Parent)
# Keep only families with _at least_ 2 offspring
ped %<>% 
     semi_join(ped %>% count(sireID,damID) %>% filter(n>1) %>% ungroup())

# GENOMIC RELATIONSHIP MATRIX
grms<-list(A=readRDS(file=here::here("output","kinship_add.rds")))

# BLUPs
blups<-readRDS(here::here("output","blups.rds")) %>%
     # need to rename the "blups" list to comply with the runCrossVal function
     rename(TrainingData=BLUPs) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     mutate(TrainingData=map(TrainingData,
                             ~filter(.,GermplasmName %in% rownames(grms$A)) %>% 
                                  # rename the germplasmName column to GID
                                  rename(GID=GermplasmName))) %>% 
     # It seems actually that runParentWiseCrossVal() wants this column named "blups"
     rename(blups=TrainingData)

# DOSAGE MATRIX
## Dosages are also needed inside the runParentWiseCrossVal() function
## Reason is that they are used to extra SNP effects from GBLUP models
dosages<-readRDS(here::here("data","dosages.rds"))

# HAPLOTYPE MATRIX
## keep only haplos for parents-in-the-pedigree
## those which will be used in prediction, saves memory
haploMat<-readRDS(file=here::here("data","haplotypes.rds"))
parents<-union(ped$sireID,ped$damID) 
parenthaps<-sort(c(paste0(parents,"_HapA"),
                   paste0(parents,"_HapB")))
haploMat<-haploMat[parenthaps,]

# SELECTION INDEX
SIwts<-c(PlantHeight=1,
         FirstBranchHeight=10,
         BranchlevelNum=-3,
         PlantArchitecture=6)
```

```{r}
# RECOMBINATION FREQUENCY MATRIX
recombFreqMat<-readRDS(file=here::here("output","recombFreqMat_1minus2c.rds"))
```
#Run cross-validation
```{r}
#Too long time
starttime<-proc.time()[3]
parentWiseCV<-runParentWiseCrossVal(nrepeats=2,nfolds=5,seed=9999999999,
                                    modelType="A",
                                    ncores=10,
                                    ped=ped,
                                    blups=blups,
                                    dosages=dosages,
                                    haploMat=haploMat,
                                    grms=grms,
                                    recombFreqMat = recombFreqMat,
                                    selInd = TRUE, SIwts = SIwts)
elapsed<-proc.time()[3]-starttime; elapsed/60
```
```{r}

```





RR-BLUP aka SNP-BLUP
```{r}
M<-readRDS(file=here::here("data","dosages.rds"))
# re-load and extract the blups for just 1 trait
blups<-readRDS(file=here::here("output","blups.rds"))
dm_blups<-blups$BLUPs[[1]]
# remove phenotyped-but-not-genoypted blups
dm_blups %<>% 
     filter(GermplasmName %in% rownames(M))
# remove genotyped-but-not-phenoypted (not strictly necessary)
M<-M[rownames(M) %in% dm_blups$GermplasmName,]
# Create a kinship matrix "K"
K<-kinship(M,"add")
# Now create a centered dosage matrix, to use as predictors for rr-BLUP
Z<-centerDosage(M)

#Fit the GBLUP model
gblup<-mmer(fixed = drgBLUP~1,
            random = ~vs(GermplasmName,Gu = K),
            weights = WT,
            data=dm_blups)
#Now fit the RRBLUP model
# one catch
# the rows of the centered dosage matrix must be matched
## to the rows of data.frame with the blups
## one way to do this is by creating and incidence matrix
## use the model.matrix() function as follows:
ZincMat=model.matrix(~factor(GermplasmName,levels=rownames(Z))-1,data=dm_blups) # -1 because we don't want an intercept here
dim(ZincMat)# [1] 3114 3114
#now fit the RRBLUP model
rrblup<-mmer(fixed = drgBLUP~1,
             random = ~vsr(list(ZincMat%*%Z),buildGu = FALSE),
             weights = WT,
             data=dm_blups)
```

