---
title: "genselect"
author: "Opamelas83"
date: "2023-04-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
```{r}
install.packages("sommer") #new version is important (why?)
library(tidyverse); library(data.table); library(genomicMateSelectR); library(magrittr); library(sommer)
```

##K-fold Cross validation
```{r}
blups<-readRDS(here::here("output","blups.rds"))
dosages<-readRDS(file=here::here("data","dosages.rds"))
A<-readRDS(file=here::here("output","kinship_add.rds"))
D<-readRDS(file=here::here("output","kinship_AD.rds"))
D<-readRDS(file=here::here("output","kinship_dom.rds"))
grms[["D"]]<-D

blups %<>% 
     # need to rename the "blups" list to comply with the runCrossVal function
     rename(TrainingData=BLUPs) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     mutate(TrainingData=map(TrainingData,
                             ~filter(.,GermplasmName %in% rownames(A)) %>% 
                                  # rename the germplasmName column to GID
                                  rename(GID=GermplasmName)))

blups
## A tibble: 4 × 2
#  Trait             TrainingData        
#  <chr>             <list>              
#1 BranchlevelNum    <tibble [3,113 × 6]>
#2 FirstBranchHeight <tibble [2,698 × 6]>
#3 PlantHeight       <tibble [3,664 × 6]>
#4 PlantArchitecture <tibble [3,470 × 6]>
#5 Cylindrical_type  <tibble [1,518 × 6]>
#6 Umbrella_type     <tibble [1,045 × 6]>
#7 Open_type         <tibble [1,164 × 6]>
#8 Compact_type      <tibble [1,518 × 6]>
#9 Erectibility      <tibble [1,045 × 6]>


#For fastest, lightest compute of accuracy, remove non-phenotyped from kinship
#gids<-blups %>% 
#     unnest(TrainingData) %$% unique(GID)
# dim(A) [1] 963 963

#D<-D[gids,gids]

```
# selection indices:To think about 
```{r}
#Include selection index weights
SIwts<-c(PlantHeight=4,
         FirstBranchHeight=4,
         BranchlevelNum=4,
         PlantArchitecture=4)
#SIwts
```

#Execute cross-validation
```{r}
#Plan to do this by breeging program too.
#For A 
standardCV<-runCrossVal(blups=blups,
                        modelType="A",
                        selInd=FALSE,
                        grms=list(A=A),
                        nrepeats=3,nfolds=5,
                        gid="GID",seed=999999999,
                        ncores=10)
#For "AD"
standardCV<-runCrossVal(blups=blups,
                        modelType="AD",
                        selInd=FALSE,
                        grms=list(A=A,D=D),
                        nrepeats=3,nfolds=5,
                        gid="GID",seed=999999999,
                        ncores=10)

#For "DirDim"
standardCV<-runCrossVal(blups=blups,
                        modelType="DirDom",
                        selInd=FALSE,
                        grms=list(A=A,D=D),
                        dosages=dosages,
                        nrepeats=3,nfolds=5,
                        gid="GID",seed=999999999,
                        ncores=8)
#standardCV %>% unnest(accuracyEstOut)
#Plot result
saveRDS(standardCV,file = here::here("output","AD_KfoldsCval.rds"))
#standardCV <- readRDS("output/AD_KfoldsCval.rds")
jpeg(filename = "output/AD_KfoldCval.jpeg")

standardCV %>% 
     unnest(accuracyEstOut) %>% 
     mutate(Trait = factor(Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture", "Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type", "Erectibility"))) %>% #
     dplyr::select(repeats,id,predOf,Trait,Accuracy) %>%
filter(!(Trait == "PlantArchitecture")) %>%
     ggplot(.,aes(x=Trait,y=Accuracy,fill=Trait)) + 
  ggtitle("AD_model") +
  geom_boxplot() + theme_bw()
dev.off()
# cross validation by breeding program
# cross validation by periode 

standardCV_A <- readRDS("output/A_KfoldsCval.rds") 
standardCV_AD <- readRDS("output/AD_KfoldsCval.rds") 
standardCV_DirDom <- readRDS("output/DirDom_KfoldsCval.rds") 

standardCV_A <- standardCV_A %>% 
     unnest(accuracyEstOut) %>% 
     mutate(Trait = factor(Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture", "Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type", "Erectibility"))) %>% #
     dplyr::select(repeats,id,predOf,Trait,Accuracy) %>%
filter(!(Trait == "PlantArchitecture")) %>%
  mutate(Model = "Additive")

standardCV_AD <- standardCV_AD %>% 
     unnest(accuracyEstOut) %>% 
     mutate(Trait = factor(Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture", "Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type", "Erectibility"))) %>% #
     dplyr::select(repeats,id,predOf,Trait,Accuracy) %>%
filter(!(Trait == "PlantArchitecture")) %>%
  mutate(Model= "Additive_Dom")
##Just to try including the dom_gen model but it did not showed any significant difference
standardCV_DirDom <- standardCV_DirDom %>% 
     unnest(accuracyEstOut) %>% 
     mutate(Trait = factor(Trait,      
                        levels = c("PlantHeight", "FirstBranchHeight", "BranchlevelNum", "PlantArchitecture", "Cylindrical_type", "Umbrella_type", "Open_type", "Compact_type", "Erectibility"))) %>% #
     dplyr::select(repeats,id,predOf,Trait,Accuracy) %>%
filter(!(Trait == "PlantArchitecture")) %>%
  mutate(Model= "Additive_Dom_Gen")

#Merge data from the tree models
standardCV_Sum <- rbind(standardCV_A,standardCV_AD) %>% filter(Trait != ("Erectibility"))# to better understanding I will replace "Erectibility" by "Erect_type"or no?
write_csv(standardCV_Sum, "Result/Genomic_standardCV_Sum.csv", append = FALSE)
#PLot results of standardCV_Sum
standardCV_Sum %>%
    ggplot(.,aes(x=Trait,y=Accuracy,fill=Model)) + 
  ggtitle("Genomic prediction") +
  geom_boxplot() + theme_bw()
dev.off()
#ANOVA for Comparing Model Accuracy for Each Trait
MyAOV <- aov(Accuracy ~ Model*Trait, data = standardCV_Sum)
summary (MyAOV)
TukeyHSD(MyAOV, which = "Trait")
Mean <- model.tables(MyAOV, type="means", se = TRUE)
Mean$tables
#rr <- standardCV_Sum %>% filter(Trait == "Compact_type")
#MyAOV <- aov(Accuracy ~ Model, data = rr)
summary (MyAOV)
# cross validation by breeding program
# cross validation by periode 

```

##Predict parental breeding values
```{r}
blups<-readRDS(here::here("output","blups.rds"))
A<-readRDS(file=here::here("output","kinship_add.rds"))

blups %<>% 
     # need to rename the "blups" list to comply with the runCrossVal function
     rename(TrainingData=BLUPs) %>% 
     dplyr::select(Trait,TrainingData) %>% 
     # need also to remove phenotyped-but-not-genotyped lines
     mutate(TrainingData=map(TrainingData,
                             ~filter(.,GermplasmName %in% rownames(A)) %>% 
                                  # rename the germplasmName column to GID
                                  rename(GID=GermplasmName)))

blups
## A tibble: 4 × 2

```


#Run genomic predictions
```{r}
gpreds<-runGenomicPredictions(modelType="A",
                              selInd=TRUE, SIwts=SIwts,
                              blups=blups,
                              grms=list(A=A),
                              ncores=3)
#Extract GEBV
gpreds
#We have a single-row tibble. To access a simple table listing GEBV for each trait and the selection index:
gpreds$gblups[[1]]

#At this point, you can use the SELIND predictions directly to rank and select parents. Example: sort by SELIND and pick the top 10?s[[1]]

gpreds$gblups[[1]] %>% 
     arrange(desc(SELIND)) %>% 
     slice(1:10)
#For more detailed output, including variance component estimates:

gpreds$genomicPredOut[[1]]

gpreds$genomicPredOut[[1]]$varcomps[[1]]
#>                         VarComp VarCompSE   Zratio

saveRDS(gpreds,file = here::here("output","genomicPredictions.rds"))
```

#Accuracy of cross prediction for training
```{r}
ped<-read_delim(here::here("data","archi_pedigree.txt"),delim = "\t")
# Keep only complete pedigree records
ped %<>% 
     dplyr::select(-Cross_Type) %>% 
     filter(!is.na(Female_Parent),
            !is.na(Male_Parent),
            Female_Parent!="?",
            Male_Parent!="?") %>% 
     distinct
# Number of full-sib families
ped %>% distinct(Female_Parent,Male_Parent) %>% nrow()
#1985
#Summarise distribution of full-sib family sizes
ped %>%   
  count(Female_Parent,Male_Parent) %>% arrange(desc(n))  %>%  summary(.$n)# We need 2 members by family for this analysis (why?)
#Full genotyped trio
dosages<-readRDS(here::here("data","dosages.rds"))
genotyped_gids<-rownames(dosages)
#Are all pedigree genotyped?
table(ped$Accession %in% genotyped_gids)
#FALSE  TRUE 
#  718  7982 
#FALSE
table(ped$Female_Parent %in% genotyped_gids)
#FALSE  TRUE 
# 1113  7587
#We need only parents and descendance present in our SNP data
genotyped_ped<-ped %>% 
     filter(Accession %in% genotyped_gids,
            Female_Parent %in% genotyped_gids,
            Male_Parent %in% genotyped_gids)
genotyped_ped %>% nrow()
# 6512 complete trios (accessions + male parent + female parents)
genotyped_ped %>% 
     count(Female_Parent,Male_Parent) %>% arrange(desc(n))  %>%  summary(.$n)
#Female_Parent      Male_Parent   looks like 1196 full-sib families.        # Length:1196        Length:1196  
genotyped_ped %>% 
     count(Female_Parent,Male_Parent) %>% 
     filter(n>1)#810 families have offsprinf >1.

```
# validate pedigree There is one additional step I highly recommend and will demonstrate here.Plant breeding pedigrees can often have errors, esp. for the male (pollen) parent. For that reason, I recommend using the genomic data to check the pedigree. We do not want our estimate of family-genetic variance prediction accuracy further detrimated by the presence of incorrect pedigree entrees.There are various software options to do this, probably an R package or two. My approach uses the --genome IBD calculator in the command-line program PLINK v1.9, click here for the PLINK1.9 manual and to download/install the program.
```{r}
#plink --vcf VCF/BreedBaseGenotypes.vcf.gz \
 #    --make-bed --const-fid --keep-allele-order \
  #   --out VCF/BreedBaseGenotypes_subset

#plink --bfile VCF/BreedBaseGenotypes_subset \
#  --genome \
 # --out ../output/BreedBaseGenotypes_subset;

```

