---
title: "Phenodata_analysis"
author: "Opamelas83"
date: "2023-03-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(tidyverse); library(data.table); library(lme4); library(here); library(magrittr); library(lubridate); library(MuMIn); library(arm); library(gt); library(corrplot); library(reactable); library(ggpubr); library(rstatix); library(purrr);
```
##Fit models to get the BLUP by trait for all the trials# Yield component have been added for testing the correlation with teh architecture traits
```{r}
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE)
#1. Continuous traits: the main analysis
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "DMC","FRWT","FTWT") #"PlantArchitecture"
ArchiTRials <- MyArchiphenos_final %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names
i <- ArchiTRaits[1]
j <- ArchiTRials[1]
results <- tibble()
for(i in ArchiTRaits){
     data1 <- MyArchiphenos_final %>% 
       filter(ArchiTRait == i)
     Data <- tibble()
  for(j in ArchiTRials){ 
     data2 <- data1 %>% 
       filter(studyName == j)
     if(nrow(data2) > 1){
       try(data2 <- data2 %>% mutate(PlantsPerPlot = ifelse(!all(is.na(PHar)), max(PHar, na.rm=T), NA),
              PropPHar = PHar / PlantsPerPlot,
              ArchiValue = ifelse(!is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = (ArchiValue / (PlantsPerPlot*PlotArea))*10,
                                  no = ifelse(is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = NA, no = ArchiValue)),
              PropPHar  = PropPHar) %>% 
          filter(!is.na(ArchiValue)))
      try(Data <- bind_rows(Data, data2)) 
      }
   }
    #Central mode
     Formula1 <- "ArchiValue ~ yearInLoc + (1|repInTrial) + (1|germplasmName)"#(1|blockInrep)
Formula <- ifelse(unique(Data$ArchiTRait) %in% c("FRWT", "FTWT") & length(unique(Data$PropPHar)) > 1, yes = paste0(Formula1," + PropPHar"), Formula1)
     try(MixedModels <- lmer(as.formula(Formula), data = Data)) #Error: vector memory exhausted (limit reached?) for germplasmName:yearInLoc/repIN in trial why we do not add GXE. 
       meanNreps <- Data  %>%
      count(germplasmName) %$% mean(n)
       varcomps <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% spread(grp,vcov)
  #     2. heritability calculation
       VarG <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2]
       #Vblock <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2]
       Vrep <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2]
       VarResidual <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2]
       H2 <- round(VarG/(VarG + ((Vrep + VarResidual)/meanNreps)),2)
   #    3. BLUP, reliability and deregressed BLUP
       BLUP <- ranef(MixedModels, condVar=TRUE)[["germplasmName"]]
      PEV <- c(attr(BLUP, "postVar"))
     try(blup<-tibble(GermplasmName=rownames(BLUP),BLUP=BLUP$`(Intercept)`,PEV=PEV) %>%
        mutate(REL=1-(PEV/VarG),
               drgBLUP=BLUP/REL,
               WT=(1-H2)/((0.1 + (1-REL)/REL)*H2)))
      try(out <- tibble(Trait = i,
                        H2 = H2,
                        VarComps = list(varcomps),
                        BLUPs = list(blup),
                        Model = Formula))
try(results <- bind_rows(results, out))
   rm(MixedModels); rm(out)
 }

saveRDS(results, file=here::here("Result","GeneralMML_result.rds"))
# COMPUTE SELIND FROM BLUPs for Genselect
blups_Archi <- results %>%
    dplyr::select(Trait,BLUPs) %>%
    filter(!Trait %in% c("DMC", "FRWT", "FTWT"))#"MCMDS"
saveRDS(blups_Archi, file=here::here("output","blups_Archi.rds"))
# COMPUTE SELIND FROM BLUPs
ArchitMML_blups <- results %>%
    dplyr::select(Trait,BLUPs) %>%
    unnest(BLUPs) %>%
    dplyr::select(GermplasmName,Trait,BLUP) %>%
    spread(Trait,BLUP) ##spread() lays out each key-value pair across multiple columns.
saveRDS(ArchitMML_blups, file=here::here("output","ArchitMML_blups.rds"))
write.table(ArchitMML_blups, file = "Result/ArchitMML_blups", quote = F,
            sep = ",", row.names = F)
#Correlation between trait using the BLUPs VALUE

BLUP <- ArchitMML_blups %>% dplyr::select(-GermplasmName) %>% dplyr::select(BranchlevelNum, FirstBranchHeight, PlantHeight, DMC, FRWT, FTWT) %>% na.omit()#PlantArchitecture, MCMDS,
BLUP_cor <- cor(BLUP)
Figplot2 <- corrplot::corrplot(BLUP_cor, method = "number", type = "upper", sig.level = 0.01, insig = "blank")
mtext("BLUP", at=6.5, line=1, cex=2)
```

## If necessary Fit models to get the BLUP by trait/breeding stage for all the trials# Yield component have been added for testing the correlation with teh architecture traits.
```{r}
MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE)
ArchiTRaits <- c("BranchlevelNum", "FirstBranchHeight", "PlantHeight", "DMC", "FRWT","FTWT")#"PlantArchitecture","MCMDS"
ArchiTRials <- MyArchiphenos_final %>% dplyr::select(studyName) %>% table %>% .[order(.)] %>% names
ArchiTRialTyp <- MyArchiphenos_final %>% dplyr::select(trialType) %>% table %>% .[order(.)] %>% names
i <- ArchiTRaits[1]
j <- ArchiTRials[1]
k <- ArchiTRialTyp[1]
results <- tibble()
for(i in ArchiTRaits){
 for(k in ArchiTRialTyp){
     data1 <- MyArchiphenos_final %>% 
       filter(trialType == k, ArchiTRait == i)
     Data <- tibble()
  for(j in ArchiTRials){ 
     data2 <- data1 %>% 
       filter(studyName == j)
     if(nrow(data2) > 1){
       try(data2 <- data2 %>% mutate(PlantsPerPlot = ifelse(!all(is.na(PHar)), max(PHar, na.rm=T), NA),
              PropPHar = PHar / PlantsPerPlot,
              ArchiValue = ifelse(!is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = (ArchiValue / (PlantsPerPlot*PlotArea))*10,
                                  no = ifelse(is.na(PlantsPerPlot) & ArchiTRait %in% c("FRWT","FTWT"), yes = NA, no = ArchiValue)),
              PropPHar  = PropPHar) %>% 
          filter(!is.na(ArchiValue)))
      try(Data <- bind_rows(Data, data2)) 
      }
   }
    Formula1 <- "ArchiValue ~ yearInLoc + (1|repInTrial) + (1|germplasmName)"#(1|blockInrep)
Formula <- ifelse(unique(Data$ArchiTRait) %in% c("FRWT", "FTWT") & length(unique(Data$PropPHar)) > 1, yes = paste0(Formula1," + PropPHar"), Formula1)
     try(MixedModels <- lmer(as.formula(Formula), data = Data)) #Error: vector memory exhausted (limit reached?) for germplasmName:yearInLoc/repIN in trial why we do not add GXE. 
       meanNreps <- Data  %>%
      count(germplasmName) %$% mean(n)
       varcomps <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% spread(grp,vcov)
       VarG <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2]
       #Vblock <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2]
       Vrep <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2]
       VarResidual <- as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[3,2]
       H2 <- round(VarG/(VarG + ((Vrep + VarResidual)/meanNreps)),2)
       BLUP <- ranef(MixedModels, condVar=TRUE)[["germplasmName"]]
      PEV <- c(attr(BLUP, "postVar"))
     try(blup<-tibble(GermplasmName=rownames(BLUP),BLUP=BLUP$`(Intercept)`,PEV=PEV) %>%
        mutate(REL=1-(PEV/VarG),
               drgBLUP=BLUP/REL,
               WT=(1-H2)/((0.1 + (1-REL)/REL)*H2)))
      try(out <- tibble(Trait = i,
                        trialType = k,
                        H2 = H2,
                        VarComps = list(varcomps),
                        BLUPs = list(blup),
                        Model = Formula))
try(results <- bind_rows(results, out))
   rm(MixedModels); rm(out)
 }
}
saveRDS(results, file=here::here("Result","trialTyMML_result.rds"))
H2trial1 <- results %>% select(Trait, trialType, H2)

# COMPUTE SELIND FROM BLUPs for Genselect
blups_Archi <- results %>%
    dplyr::select(Trait,BLUPs) %>%
    filter(!Trait %in% c("DMC", "FRWT", "FTWT", "MCMDS"))
saveRDS(blups_Archi, file=here::here("output","blups_Archi.rds"))
# COMPUTE SELIND FROM BLUPs
ArchitMML_blups <- results %>%
    dplyr::select(Trait,BLUPs) %>%
    unnest(BLUPs) %>%
    dplyr::select(GermplasmName,Trait,BLUP) %>%
    spread(Trait,BLUP) ##spread() lays out each key-value pair across multiple columns.
saveRDS(ArchitMML_blups, file=here::here("output","ArchitMML_blups.rds"))
write.table(ArchitMML_blups, file = "Result/ArchitMML_blups", quote = F,
            sep = ",", row.names = F)
#Correlation between trait using the BLUPs VALUE

BLUP <- ArchitMML_blups %>% dplyr::select(-GermplasmName) %>% dplyr::select(BranchlevelNum, FirstBranchHeight, PlantHeight, DMC, FRWT, FTWT) %>% na.omit()#PlantArchitecture, MCMDS,
BLUP_cor <- cor(BLUP)
Figplot2 <- corrplot::corrplot(BLUP_cor, method = "number", type = "upper", sig.level = 0.01, insig = "blank")
mtext("BLUP", at=6.5, line=1, cex=2)
```

# For Shape trait:Plant-shape traits were treated differently
```{r}
Shapephenos_filtered <- read_csv("data/Shapephenos_filtered.csv", col_names = TRUE)
Shapephenos_filtered <- Shapephenos_filtered %>% filter(!ShapeTRait == "Erectibility")
Shapephenos_filtered$ShapeTRait <- 
  fct_recode(Shapephenos_filtered$ShapeTRait, 
             "Cylindrical" = "Cylindrical_type", 
             "Open" = "Open_type",
             "Compact" = "Compact_type", 
             "Umbrella" = "Umbrella_type")

ShapeTRaits <- c("Cylindrical", "Umbrella", "Open", "Compact") #"Erectibility"

#recoding the shape trait to facilitate the correlation interpretation 
Shapephenos_filtered$ShapeValue = ifelse(Shapephenos_filtered$ShapeValue %in% str_subset(Shapephenos_filtered$ShapeValue, "Non"), 0, 1)
i <- ShapeTRaits [1]
results_Shape<- tibble()
for(i in ShapeTRaits){
     Data <- Shapephenos_filtered %>% 
       filter(ShapeTRait == i) %>%
       
     filter(!is.na(germplasmName))
     if(nrow(Data) > 1){
      if(length(unique(Data$ShapeValue)) > 1){ 
     try(MixedModels_Shape <- lme4::glmer(ShapeValue ~ yearInLoc + (1|repInTrial) + (1|germplasmName), data = Data,  family = binomial, nAGQ = 0))#nAGQ Values greater produce greater accuracy but produce error message in my case.#(1|blockInrep) effect is almost 0
                     meanNreps <- Data  %>%
      count(germplasmName) %$% mean(n)
       varcomps <- as.data.frame(VarCorr(MixedModels_Shape))[,c("grp","vcov")] %>% spread(grp,vcov)
       varcomps$Residual <- as.data.frame(var(residuals(MixedModels_Shape))) %>% .[1,1] 
       VarG <- as.data.frame(VarCorr(MixedModels_Shape))[,c("grp","vcov")] %>% .[1,2]
       Vrep <- as.data.frame(VarCorr(MixedModels_Shape))[,c("grp","vcov")] %>% .[2,2]
       VarResidual <- as.data.frame(var(residuals(MixedModels_Shape))) %>% .[1,1]
       H2 <- round(VarG/(VarG + ((Vrep + VarResidual)/meanNreps)),2)
       BLUP <- ranef(MixedModels_Shape, condVar=TRUE)[["germplasmName"]]
      PEV <- c(attr(BLUP, "postVar"))
     try(blup<-tibble(GermplasmName=rownames(BLUP),BLUP=BLUP$`(Intercept)`,PEV=PEV) %>%
        mutate(REL=1-(PEV/VarG),
               drgBLUP=BLUP/REL,
               WT=(1-H2)/((0.1 + (1-REL)/REL)*H2)))
      try(out <- tibble(Trait = i,
                        H2 = H2,
                        VarComps = list(varcomps),
                        BLUPs = list(blup)))
try(results_Shape <- bind_rows(results_Shape, out))
   #rm(MixedModels_Shape); rm(out)
      }
  }   
}

saveRDS(results_Shape, file=here::here("Result","results_Shape.rds"))
# COMPUTE SELIND FROM BLUPs for Genselect
blups_Shape <- results_Shape %>%
    dplyr::select(Trait,BLUPs)
saveRDS(blups_Shape, file=here::here("output","blups_shape.rds"))
# COMPUTE SELIND FROM BLUPs
ShapeMML_blups <- results_Shape %>%
    dplyr::select(Trait,BLUPs) %>%
    unnest(BLUPs) %>%
    dplyr::select(GermplasmName,Trait,BLUP) %>%
    spread(Trait,BLUP) ##spread() lays out each key-value pair across multiple columns.
saveRDS(ShapeMML_blups, file=here::here("output","ShapeMML_blups.rds"))
write.table(ShapeMML_blups, file = "Result/ShapeMML_blups", quote = F,
            sep = ",", row.names = F)
#Correlation between trait using the BLUPs VALUE
ArchitMML_blups <- readRDS("output/ArchitMML_blups.rds")
ShapeMML_blups <- readRDS("output/ShapeMML_blups.rds")
ShapeArchiMML_blups <- left_join(ShapeMML_blups, ArchitMML_blups)
ShapeMML_blups <- ShapeArchiMML_blups %>% dplyr::select(-GermplasmName) %>% dplyr::select(BranchlevelNum, FirstBranchHeight, PlantHeight, Cylindrical, Umbrella, Open, Compact)  %>% na.omit() #, MCMDS, DMC, FRWT, FTWT,PlantArchitecture,Erectibility)
ShapeMML_blups_cor <- cor(ShapeMML_blups)
Figplot2 <- corrplot::corrplot(ShapeMML_blups_cor, method = "number", type = "upper", sig.level = 0.01, insig = "blank")
mtext("BLUP", at=8.5, line=2.5, cex=2)
```
#Merge both two groups of blups to mame one file for genomic selection
```{r}
blups_Archi <- readRDS("output/blups_Archi.rds")
blups_Shape <- readRDS("output/blups_Shape.rds")
blups <- bind_rows(blups_Archi,blups_Shape)
saveRDS(blups, file=here::here("output","blups.rds"))
```
