---
title: "VCFfilestreatment"
author: "Opamelas83"
date: "2023-04-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

##The genotype data have been dowloading under 10 separated files because of the size. I should joint them to make one big file ref "https://wolfemd.github.io/GenomicSelectionManual/intro-to-genomic-prediction.html"
```{r}
library(tidyverse); library(data.table); library(vcfR); library(genomicMateSelectR); library(magrittr); library(factoextra); library(FactoMineR)
```

```{r}
#I cut the original vcf file in 2 parts to make easier the download.
#split -l 5500 data/VCF/Architect_acces_final.txt data/VCF/Architect_acces_ 

##mo367@AG-PBG-92GML7L ~ % cd Desktop/Analy_project/Plantarchitect/ 
#vcftools --vcf data/BreedBaseGenotypesDownload_a.vcf  
#VCFtools - 0.1.16
#(C) Adam Auton and Anthony Marcketta 2009
#Parameters as interpreted:
#	--vcf data/BreedBaseGenotypesDownload_a.vcf
#After filtering, kept 1165 out of 1165 Individuals
#After filtering, kept 61239 out of a possible 61239 Sites
#Run Time = 5.00 seconds
#Check duplicate
#bcftools query --list-samples data/BreedBaseGenotypes_a.vcf 
#Duplicated sample name 'IITA-TMS-IBA000070'
#Write just the column-names of the VCF to disk.
#egrep "^#CHROM" data/BreedBaseGenotypesDownload_a.vcf | head -n1 > data/vcf_colnames.txt
```

```{r}
#After download vcf file from Cassavabase (10 subseted vcf file; a,b,c...)
vcf_sample_names<-readLines("data/vcf_colnames.txt") %>% 
     strsplit(.,"\t") %>% unlist() %>% 
     .[10:length(.)]
# Check how many sample names are duplicated?
#table(duplicated(vcf_sample_names))
#1100    65 
```

```{r}
# create unique names for each VCF
unique_names_for_vcf<-tibble(vcfName=vcf_sample_names) %>% 
     # create an overall index to ensure I can recover the original column order
     mutate(vcfIndex=1:n()) %>% 
     # now for each vcfName create an sub-index, to distinguish among duplicates
     group_by(vcfName) %>% 
     # sub-index
     mutate(vcfNameIndex=1:n(),
            # For the first (or only) instance of each unique vcfName
            vcfName_Unique=ifelse(vcfNameIndex==1,
                                  # return the original name
                                  vcfName,
                                  # for all subsequent (duplicate) names, 
                                  #put a unique-ified name by pasting the sub-index
                                  paste0(vcfName,".",vcfNameIndex)))
# Write the "unique_names_for_vcf.txt" to disk
write.table(unique_names_for_vcf$vcfName_Unique,file = "data/unique_names_for_vcf.txt",
            row.names = F, col.names = F, quote = F)
# Create also a list containing only one instance of each unique name, the first instance 
subset_unique_names_for_vcf<-unique_names_for_vcf %>% 
     filter(vcfNameIndex==1) %$%
     vcfName_Unique
# Write that list to disk for subsetting the VCF downstream
write.table(subset_unique_names_for_vcf,file = "data/subset_unique_names_for_vcf.txt",
            row.names = F, col.names = F, quote = F)
```

```{r}
# replace sample names in original VCF with unique ones (creates a new VCF)
#bcftools reheader --samples data/unique_names_for_vcf.txt data/VCF/BreedBaseGenotypesDownload_a.vcf > data/BreedBaseGenotypesDownload_1.vcf;
# overwrite the original VCF with the new  that has unique names
#mv data/BreedBaseGenotypesDownload_1.vcf data/VCF/BreedBaseGenotypesDownload_a.vcf;
# check that the names are now unique by printing sample list
#Plantarchitect % bcftools query --list-samples data/VCF/BreedBaseGenotypesDownload_a.vcf
#vcftools --vcf data/VCF/BreedBaseGenotypesDownload_a.vcf --keep data/subset_unique_names_for_vcf.txt --recode --stdout | bgzip -c > data/VCF/BreedBaseGenotypes_subset_a.vcf.gz
####more
mo367@AG-PBG-92GML7L Plantarchitect_PLINK % vcftools --vcf BreedBaseGenotypes.vcf.gz --keep Myindiv.txt --recode --stdout | bgzip -c > BreedBaseGenotypesM.vcf.gz
mo367@AG-PBG-92GML7L Plantarchitect_PLINK % vcftools --vcf BreedBaseGenotypes.vcf.gz --keep Myindiv.txt --recode --stdout | bgzip -c > BreedBaseGenotypesM.vcf.gz
mo367@AG-PBG-92GML7L Plantarchitect_PLINK % vcftools --vcf BreedBaseGenotypes.vcf.gz --keep Myindiv.txt --recodemo367@AG-PBG-92GML7L Plantarchitect_PLINK % vcftools --gzvcf BreedBaseGenotypes.vcf.gz --keep Myindiv.txt --recode --stdout | bgzip -c > BreedBaseGenotypesM.vcf.gz
mo367@AG-PBG-92GML7L Plantarchitect_PLINK % vcftools --gzvcf BreedBaseGenotypes.vcf.gz --keep Myindiv.txt --recode --stdout | bgzip -c > BreedBaseGenotypesM.vcf.gz
#mo367@AG-PBG-92GML7L Plantarchitect_PLINK % bcftools query -l BreedBaseGenotypesM.vcf.gz > vcf_colnames.txt  
Myheader <- readLines("/Users/mo367/Desktop/Plantarchitect_PLINK/vcf_colnames.txt")
Myheader1 <- tibble(Myheader)
Myheader1$Myheader <- gsub(pattern = "-", replacement = "", x = Myheader1$Myheader)
Myheader1$Myheader <- gsub(pattern = " ", replacement = "", x = Myheader1$Myheader)
Myheader1$Myheader <- gsub(pattern = "[[:punct:]]", replacement = "", x = Myheader1$Myheader)
write.table(Myheader1$Myheader, file = "/Users/mo367/Desktop/Plantarchitect_PLINK/Myheader.txt",
            row.names = F, col.names = F, quote = F)
```

Shell command
```{r}
##shell for merging my vcf files. 
#bcftools index BreedBaseGenotypes_subset_a.vcf.gz && bcftools index BreedBaseGenotypes_subset_b.vcf.gz && bcftools index BreedBaseGenotypes_subset_c.vcf.gz && bcftools index BreedBaseGenotypes_subset_d.vcf.gz && bcftools index BreedBaseGenotypes_subset_e.vcf.gz && bcftools index BreedBaseGenotypes_subset_f.vcf.gz && bcftools index BreedBaseGenotypes_subset_g.vcf.gz && bcftools index BreedBaseGenotypes_subset_h.vcf.gz && bcftools index BreedBaseGenotypes_subset_i.vcf.gz && bcftools index BreedBaseGenotypes_subset_j.vcf.gz
##Then run 
#bcftools merge BreedBaseGenotypes_subset_a.vcf.gz BreedBaseGenotypes_subset_b.vcf.gz BreedBaseGenotypes_subset_c.vcf.gz  BreedBaseGenotypes_subset_d.vcf.gz BreedBaseGenotypes_subset_e.vcf.gz BreedBaseGenotypes_subset_f.vcf.gz BreedBaseGenotypes_subset_g.vcf.gz BreedBaseGenotypes_subset_h.vcf.gz BreedBaseGenotypes_subset_i.vcf.gz BreedBaseGenotypes_subset_j.vcf.gz > BreedBaseGenotypes.vcf.gz
# removing a list of genotypes 
#vcftools --gzvcf data/VCF/BreedBaseGenotypes.vcf.gz --keep data/Architect_acces_final.txt --recode --stdout | bgzip -c > data/VCF/BreedBaseGenotypes_subset.vcf.gz(txt is the list of my genotype)
#vcftools --gzvcf data/VCF/BreedBaseGenotypes_subset.vcf.gz
```
#For my training
```{r}
#mo367@AG-PBG-92GML7L Plantarchitect % bcftools query --list-samples data/BreedBaseGenotypesDownload_a.vcf
#convert to a vcf.gz format, because the file is not a real vcf file
# Plantarchitect % bcftools view -O z -o data/BreedBaseGenotypes_subset.vcf.gz data/BreedBaseGenotypesDownload_a.vcf 
```

```{r}
#Extract haplotypes from VCF with bcftools convert --hapsample
#bcftools convert --hapsample data/BreedBaseGenotypes_subset data/BreedBaseGenotypes_subset.vcf.gz
#Hap file: data/BreedBaseGenotypes_subset.hap.gz
#Sample file: data/BreedBaseGenotypes_subset.samples
#[W::vcf_parse_format] FORMAT 'NT' at 1:21745 is not defined in the header, assuming Type=String
#61239 records written, 0 skipped: 0/0/0 no-ALT/non-biallelic/filtered
```

```{r}
phenos <- read_csv("data/MyArchiphenos_final.csv")
germplasm_with_phenos<-unique(phenos$germplasmName)
subset_unique_names_for_vcf<-read.table(file = "data/subset_unique_names_for_vcf.txt", stringsAsFactors = F)$V1
table(germplasm_with_phenos %in% subset_unique_names_for_vcf)
#FALSE  TRUE 
#12080  5933
#geno and pheno
subset_unique_names_for_vcf[subset_unique_names_for_vcf %in% germplasm_with_phenos]
#5933
#pheno not geno
germplasm_with_phenos[!germplasm_with_phenos %in% subset_unique_names_for_vcf]
#[12080]
subset_unique_names_for_vcf[!subset_unique_names_for_vcf %in% germplasm_with_phenos]
#[4924]
```
#Filter VCF file for keeping geno with pheno
```{r}
#egrep "^#CHROM" data/VCF/BreedBaseGenotypes.vcf.gz | head -n1 > data/VCF/vcf_colnamesT.txt
vcf_sample_names<-readLines("data/vcf_colnamesT.txt") %>% 
     strsplit(.,"\t") %>% unlist() %>% 
     .[10:length(.)]
vcf_names <- tibble(germplasmName=vcf_sample_names)
phenos <- read_csv("data/MyArchiphenos_final.csv")
phenos_germ <- phenos %>% dplyr::select(germplasmName)
shapephenos <- read_csv("data/Shapephenos_filtered.csv")
shapephenos_germ <- shapephenos %>% dplyr::select(germplasmName)
totalphenos <- bind_rows(phenos_germ,shapephenos_germ)
phenosID <- tibble(unique(totalphenos))
#make my list for further analysis
germplasm_with_pheno <- inner_join(vcf_names, phenosID)
#germplasm_with_pheno$germplasmName# 7417 
write.table(germplasm_with_pheno$germplasmName,file = "data/germplasm_with_pheno.txt",
            row.names = F, col.names = F, quote = F)
#vcftools --gzvcf data/VCF_old/BreedBaseGenotypes.vcf.gz --keep data/germplasm_with_pheno.txt --recode --stdout | bgzip -c > data/VCF/BreedBaseGenotypes.vcf.gz 
```

#Haplotype and dosage matrix from VCF
```{r}
##Shell: Extract haplotypes from VCF with bcftools
#bcftools convert --hapsample data/VCF/BreedBaseGenotypes_subset data/VCF/BreedBaseGenotypes.vcf.gz
#Hap file: data/VCF/BreedBaseGenotypes_subset.hap.gz
#Sample file: data/VCF/BreedBaseGenotypes_subset.samples
#61239 records written, 0 skipped: 0/0/0 no-ALT/non-biallelic/filtered
library("R.utils"); library(genomicMateSelectR); library(data.table)
#Read haps to R and format them.
vcfName<-"BreedBaseGenotypes_subset"
 haps<-fread(paste0("data/VCF/",vcfName,".hap.gz"),
            stringsAsFactors = F,header = F) %>% 
  as.data.frame
 sampleids<-fread(paste0("data/VCF/",vcfName,".samples"),
                 stringsAsFactors = F,header = F,skip = 2) %>% 
  as.data.frame
hapids<-sampleids %>% 
     dplyr::select(V1,V2) %>% 
     mutate(SampleIndex=1:nrow(.)) %>% 
     rename(HapA=V1,HapB=V2) %>% 
     pivot_longer(cols=c(HapA,HapB),
                  names_to = "Haplo",values_to = "SampleID") %>% 
     mutate(HapID=paste0(SampleID,"_",Haplo)) %>% 
     arrange(SampleIndex)
colnames(haps)<-c("Chr","HAP_ID","Pos","REF","ALT",hapids$HapID)
dim(haps)
#61239 21721
haps %<>% 
     mutate(HAP_ID=gsub(":","_",HAP_ID)) %>% 
     column_to_rownames(var = "HAP_ID") %>% 
     dplyr::select(-Chr,-Pos,-REF,-ALT) %>% 
     t(.) %>% 
     as.matrix(.)
dim(haps)
# 21716 61239
##The counted allele in the dosage matrix, which will be used downstream to construct kinship matrices and estimate marker effects, should be the ALT allele. This will ensure a match to the haplotype matrix, where a “1” indicates the presence of the ALT allele.Here’s atidyverse-based approach, using group_by() plus summarise() to sum the two haplotypes for each individual across all loci.(From Marnin)

dosages<-haps %>%
     as.data.frame(.) %>% 
     rownames_to_column(var = "GID") %>% 
     separate(GID,c("SampleID","Haplo"),"_Hap",remove = T) %>% 
     dplyr::select(-Haplo) %>% 
     group_by(SampleID) %>% 
     summarise(across(everything(),~sum(.))) %>% 
     ungroup() %>% 
     column_to_rownames(var = "SampleID") %>% 
     as.matrix %>% 
     # preserve same order as in haps
     .[sampleids$V1,]

dim(dosages)
#[1] 10858 61239
```

Variant filters
```{r}
# In this case, simple: keep only positions with >1% minor allele frequency.
dosages<-maf_filter(dosages,thresh = 0.01)
dim(dosages)
#[1] 10858 61158
haps<-haps[,colnames(dosages) ]

saveRDS(dosages,file=here::here("data","dosages.rds"))
saveRDS(haps,file=here::here("data","haplotypes.rds"))
```

#Genomic Relationship matrices (GRMs)#genomicMateSelectR function kinship() to construct additive (A) and dominance (D) relationship matrices.
```{r}
A<-kinship(dosages,type="add")
D<-kinship(dosages,type="domGenotypic")
D<-kinship(dosages,type="domClassic")

 saveRDS(A,file=here::here("output","kinship_add.rds"))
 saveRDS(D,file=here::here("output","kinship_dom.rds"))
 saveRDS(D,file=here::here("output","kinship_AD.rds"))
# In this case, simple: keep only positions with >1% minor allele frequency.
```

#Recombination Frequency Matrix: source a single genome-wide file representing the ICGMC concensus genetic map on the V6 Cassava Reference genome. The file is on the Cassavabase FTP-archive
```{r}
genmap<-read.table("https://cassavabase.org/ftp/marnin_datasets/NGC_BigData/CassavaGeneticMap/cassava_cM_pred.v6.allchr.txt",
                   header = F, sep=';', stringsAsFactors = F) %>% 
     rename(SNP_ID=V1,Pos=V2,cM=V3) %>% 
  as_tibble
genmap %>% dim
#[1] 120979      3

snps_genmap<-tibble(DoseSNP_ID=colnames(dosages)) %>% 
     separate(DoseSNP_ID,c("Chr","Pos","Ref","Alt"),remove = F) %>% 
     mutate(SNP_ID=paste0("S",Chr,"_",Pos)) %>% 
     full_join(genmap %>% 
                    separate(SNP_ID,c("Chr","POS"),"_",remove = F) %>% 
                   dplyr::select(-POS) %>% 
                    mutate(Chr=gsub("S","",Chr)) %>% 
                    mutate(across(everything(),as.character)))
 library(ggplot2)
jpeg(filename = "output/snp_genmap_map.jpeg") 
snps_genmap %>% 
  ggplot(.,aes(x=as.integer(Pos)/1000/1000,y=as.numeric(cM))) +
  geom_point() +
  theme_bw() +
  facet_wrap(~as.integer(Chr))
dev.off()

```
#Interpolate genetic map
```{r}
interpolate_genmap<-function(data){
+ data_forspline<-data %>% 
    filter(!is.na(cM)) %>% 
    mutate(cumMax=cummax(cM),
           cumIncrement=cM-cumMax) %>% 
    filter(cumIncrement>=0) %>% 
    dplyr::select(-cumMax,-cumIncrement)
  
  spline<-data_forspline %$% smooth.spline(x=Pos,y=cM,spar = 0.75)
  
  splinemap<-predict(spline,x = data$Pos) %>% 
    as_tibble(.) %>% 
    rename(Pos=x,cM=y) %>% 
    mutate(cumMax=cummax(cM),
           cumIncrement=cM-cumMax) %>% 
    mutate(cM=cumMax) %>% 
    dplyr::select(-cumMax,-cumIncrement)
  
  return(splinemap) 
}
 splined_snps_genmap<-snps_genmap %>% 
  dplyr::select(-cM) %>% 
  mutate(Pos=as.numeric(Pos)) %>% 
  left_join(snps_genmap %>% 
              mutate(across(c(Pos,cM),as.numeric)) %>% 
              arrange(Chr,Pos) %>% 
              nest(-Chr) %>% 
              mutate(data=map(data,interpolate_genmap)) %>% 
              unnest(data)) %>% 
  distinct
all(colnames(dosages) %in% splined_snps_genmap$DoseSNP_ID)

```

 splined map
```{r}
jpeg(filename = "output/splined_map.jpeg")
splined_snps_genmap %>% 
     filter(DoseSNP_ID %in% colnames(dosages)) %>% 
     mutate(Map="Spline") %>% 
     bind_rows(snps_genmap %>% 
                    filter(DoseSNP_ID %in% colnames(dosages),
                           !is.na(cM)) %>% 
                    mutate(across(c(Pos,cM),as.numeric)) %>% 
                    arrange(Chr,Pos) %>% mutate(Map="Data")) %>% 
  ggplot(.,aes(x=Pos/1000/1000,y=cM,color=Map, shape=Map),alpha=0.3,size=0.75) + 
  geom_point() + 
  theme_bw() + facet_wrap(~as.integer(Chr), scales='free_x')
 dev.off()
 
splined_snps_genmap %>% 
     filter(DoseSNP_ID %in% colnames(dosages)) %>% 
     saveRDS(.,file=here::here("output","interpolated_genmap.rds"))
#Recomb. freq. matrix
genmap<-readRDS(file=here::here("output","interpolated_genmap.rds"))
m<-genmap$cM;
names(m)<-genmap$DoseSNP_ID
recombFreqMat<-1-(2*genmap2recombfreq(m,nChr = 18))
saveRDS(recombFreqMat,file=here::here("output","recombFreqMat_1minus2c.rds"))

```

#Histogram of kinship matrix
```{r}
##call the Kinship matrix. The matrix was constructed using the same formula implemented in the “classic” rrBLUP function A.mat(): From VanRaden (2008), Method 1.
A<-readRDS(file=here::here("output","kinship_add.rds"))
#The off-diagonals express the genomic relatedness between different individuals in the dataset.
#This formulation has a mean-value of zero, such that >0 kinship means “more than the average population level of relatedness” while <0 values mean “greater than average.”
jpeg(filename = "output/kinship_offdiag.jpeg")
hist(A[lower.tri(A)], main='Off-diagonal values of the kinship matrix', xlim = c(-0.5,0.7))
dev.off()
#The diagonal values express the relatedness of each individual to itself, which should be interpreted as the level of inbreeding (homozygosity).
jpeg(filename = "output/kinship_diag.jpeg")
hist(diag(A), main='Diagonals of the kinship', breaks=20)
dev.off()
```

```{r}
#Mydosages <- readRDS("/workdir/Plantarchitect/data/dosages.rds")
#Mydosages_ID <- rownames(Mydosages) %>% as_tibble() %>% mutate(germplasmName = value) %>% dplyr::select(germplasmName)
#MyArchiphenos_final <- read_csv("data/MyArchiphenos_final.csv", col_names = TRUE)
#Mydosages_ID_Prog <- MyArchiphenos_final %>% group_by(germplasmName, programName) %>% summarise(programName = unique(programName))
#Mydosages_IDprog <- left_join(Mydosages_ID, Mydosages_ID_Prog)
#Mydosages
#Germplas_program <- MyArchiphenos %>% dplyr::select(germplasmName, programName) 
#res.pca <- prcomp(Mydosages) [FactoMineR package] My#
```
